### 2022-02-22 notes/notes.tex
Some editing

\documentclass{article}
\usepackage{hyperref}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}

\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{amsfonts}
\usepackage{amssymb}

\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{tikz}
\usepackage[capitalize]{cleveref}

\usetikzlibrary{fit,positioning,trees,cd}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}

\crefname{algocf}{algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}

\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{observation}{Observation}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\makeatletter
\newcommand\incircbin
    {%
      \mathpalette\@incircbin
    }
    \newcommand\@incircbin[2]
                          {%
                            \mathbin%
                                {%
                                  \ooalign{\hidewidth$#1#2$\hidewidth\crcr$#1\bigcirc$}%
                                }%
                          }
                          \newcommand{\oland}{\incircbin{\land}}
                          \newcommand{\olor}{\incircbin{\lor}}
                          \newcommand{\Contradiction}{\incircbin{\bot}}
                          \newcommand{\Tautology}{\incircbin{\top}}
                          \newcommand{\Smoothing}{\incircbin{}}
                          \newcommand{\Unit}{\incircbin{1}}
                          \makeatother

\newcommand\pfun{\mathrel{\ooalign{\hfil$\mapstochar\mkern5mu$\hfil\cr$\to$\cr}}}
\newcommand\mdoubleplus{\mathbin{+\mkern-10mu+}}
\newcommand*{\twoheadrightarrowtail}{\mathrel{\rightarrowtail\kern-1.9ex\twoheadrightarrow}}

\DeclareMathOperator{\CR}{\textsc{CR}}
\DeclareMathOperator{\DR}{\textsc{DR}}
\DeclareMathOperator{\IE}{\textsc{IE}}
\DeclareMathOperator{\Reff}{\textsc{Ref}}
\DeclareMathOperator{\wwp}{w}
\DeclareMathOperator{\wwn}{\overline{w}}

\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\Imm}{Im}
\DeclareMathOperator{\Doms}{Doms}
\DeclareMathOperator{\size}{\sigma}
\DeclareMathOperator{\Vars}{Vars}
\DeclareMathOperator{\WMC}{WMC}
\DeclareMathOperator{\gr}{gr}

\title{Recursive Solutions to First-Order Model Counting}

\begin{document}
\maketitle

\section{Definitions}

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[triangle/.style = {regular polygon, regular polygon sides=3}]
    \node[draw,ellipse] (a) at (0, 0) {$\phi(n)$};
    \node[draw,ellipse] (b) at (0, -2) {$\psi(n-1)$};
    \node[draw,ellipse] (c) at (-1, -4) {$\phi(n-1)$};
    \node[draw,triangle] (d) at (1, -4) {};
    \draw[-{Latex[length=2mm]}] (a) -- (b) node [midway,xshift=50] {domain recursion};
    \draw[-{Latex[length=2mm]}] (b) -- (c);
    \draw[-{Latex[length=2mm]}] (b) -- (d);
    \draw[-{Latex[length=2mm]}] (c) to [bend left=45] node [midway,xshift=-30] () {$n \mapsto n-1$} (a);
  \end{tikzpicture}
  \caption{An illustration of the main idea. TODO: refer to this in the introduction.}
  \label{fig:idea}
\end{figure}

\paragraph{Things I might need to explain.}
\begin{itemize}
\item atom, (positive/negative) literal, constant, predicate, variable, literal variable, clause, unit clause
\item $\Vars$, $\Vars(c) = \Vars(L) \cup \Vars(C)$
\item $\Doms$ on both formulas and clauses. $\Doms(c) = \Imm \delta_c$, and $\Doms(\phi) = \bigcup_{c \in \phi} \Doms(c)$.
\item $\WMC$, $\wwp$, $\wwn$, $\Imm$
\item two parts: compilation and inference.
\item During inference, there is a domain size map $\size\colon \mathcal{D} \to \mathbb{N}_0$.
\item mention which rules are in $\Gamma$ and which ones are in $\Delta$ (and why tryCache has to be in $\Delta$).
\item \textsc{ForcLift}
\item In a way, we're dividing the idea of domain recursion between the IDR and the Ref nodes, thus also generalising it. [This is good context to refer to \cref{fig:idea}.]
\end{itemize}

\paragraph{Notation.}
\begin{itemize}
\item We write $\to$ for functions, $\pfun$ for partial functions, $\twoheadrightarrowtail$ for bijections, and $\hookrightarrow$ for set inclusion.
\end{itemize}

\paragraph{TODO}
\begin{itemize}
\item capitalise variable names.
\item use $\langle\rangle$ for lists. Square brackets are overloaded already.
\end{itemize}

Most of the definitions here are adaptations/formalisations of \cite{DBLP:conf/ijcai/BroeckTMDR11} and the corresponding code.

\begin{definition}
  A \emph{domain} is a symbol for a finite set.\footnote{In the context of functions, the domain of a function $f$ retains its usual meaning and is denoted $\dom(f)$.} Let $\mathcal{D}$ be the set of all domains. Note that this set expands during the compilation.
\end{definition}

\begin{definition}
An \emph{(inequality) constraint} is a pair $(a, b)$, where $a$ is a variable, and $b$ is either a variable or a constant.
\end{definition}

\begin{definition}
  A \emph{clause} is a triple $c = (L, C, \delta_c)$, where $L$ is the set of literals, $C$ is a set of constraints, and $\delta_c\colon \Vars(c) \to \mathcal{D}$ is a function that maps all variables in $c$ to their domains such that (s.t.) if $(x, y) \in C$ for some variables $x$ and $y$, then $\delta_c(x) = \delta_c(y)$. Equality of clauses is defined in the usual way (i.e., all variables, domains, etc. must match). TODO: we will always use this subscript notation for the $\delta$'s. 
\end{definition}

A \emph{formula} is a set of clauses.

We use hash codes to efficiently check whether a recursive relationship between two formulas is plausible. (It is plausible if the formulas are equal up to variables and domains.) The hash code of a clause $c = (L, C, \delta_c)$ combines the hash codes of the sets of constants and predicates in $c$, the numbers of positive and negative literals, the number of inequality constraints $|C|$, and the number of variables $|\Vars(c)|$. The hash code of a formula $\phi$ combines the hash codes of all its clauses and is denoted $\#\phi$.

\begin{definition}
  Let $\gr(\cdot; \sigma)$ be the function (parameterised by the domain size function $\size$) that takes a clause $c = (L, C, \delta)$ and returns the number of ways the variables in $c$ can be replaced by elements of their respective domains in a way that satisfies the inequality constraints.\footnote{Note that the literals of the clause have no effect on $\gr$.} Formally, for each variable $v \in \Vars(c)$, let $C_v = \{ w \mid (u, w) \in C \setminus \Vars(c)^2 \text{, }u \ne v \}$ be the set of (explicitly named) constants that $v$ is permitted to be equal to. Then
  \[
  \gr(c; \sigma) \coloneqq \left|\left\{\, (e_v)_{v \in \Vars(c)} \in \prod_{v \in \Vars(c)} C_v \sqcup [\sigma(\delta(v)) - |C_v|] \;\middle|\; \right.\right. e_u \ne e_w \text{ for all } (u, w) \in C \cap \Vars(c)^2 \left.\left. \vphantom{\prod_{v \in \Vars(c)}} \,\right\}\right|
  \]
  for any clause $c$. (Here, $[n] \coloneqq \{\,1, 2, \dots, n\,\}$ for any non-negative integer $n$.)
\end{definition}

TODO: how does the algorithm prevent the number in $[\cdot]$ from being negative?

\paragraph{Notation for lists.}
We let $[]$ and $[x]$ denote an empty list and a list with one element $x$, respectively. We write $\in$ for (in-order) enumeration, $\mdoubleplus$ for concatenation, and $|\cdot|$ for the length of a list. Let $h : t$ denote a list with first element (a.k.a. head) $h$ and remaining list (a.k.a. tail) $t$. We also use list comprehensions written equivalently to set comprehensions. For example, let $L \coloneqq [1]$ and $M \coloneqq [2]$ be two lists. Then $M = [2x \mid x \in L]$, $L \mdoubleplus M = 1 : [2]$, and $|M| = 1$.

\section{Search/Compilation}

\subsection{From Circuits to Labelled Graphs}

A \emph{first-order deterministic decomposable negation normal form computational graph} (FCG) is a (weakly connected) directed graph with a a single source, vertex labels, and ordered outgoing edges.\footnote{Note that imposing an ordering on outgoing edges is just a limited version of edge labelling.} We denote an FCG as $G = (V, s, N^+, \tau)$, where $V$ is the set of vertices, and $s \in V$ is the unique source. Also, $N^+$ is the direct successor function that maps each vertex in $V$ to a \emph{list} that contains either other vertices in $V$ or a special symbol $\star$. This symbol means that the target of the edge is yet to be determined.

Vertex labels consist of two parts: the \emph{type} and the \emph{parameters}. For the main definition, we leave the parameters implicit and let $\tau\colon V \to \mathcal{T}$ denote the vertex-labelling function that maps each vertex in $V$ to its type in $\mathcal{T}$. Most of our list of types $\mathcal{T} \coloneqq \{\, \Smoothing, \Unit, \Tautology, \Contradiction, \oland, \olor, \CR, \DR, \IE, \Reff \,\}$ is as described in previous work \cite{DBLP:conf/nips/Broeck11,DBLP:conf/ijcai/BroeckTMDR11} as well as the source code of \textsc{ForcLift}\footnote{\url{https://dtai.cs.kuleuven.be/drupal/wfomc}} but with one new type $\CR$ and two revamped types $\DR$ and $\Reff$. For each vertex $v \in V$, the length of the list $N^+(v)$ (i.e., the out-degree of $v$) is determined by its type $\tau(v)$.

As in previous work \cite{DBLP:conf/ijcai/BroeckTMDR11}, to describe individual vertices and small (sub)-FCGs, we also use notation of the form, e.g., $\Reff_\rho(v)$. Here, the type of the vertex (e.g., $\Reff$) is `applied' to its direct successors (e.g., $v$) using either function or infix notation and provided with its parameter(s) (e.g., $\rho$) in the subscript. We say that `$G$ is an FCG for formula $\phi$' if two conditions are satisfied. First, the image of $N^+$ contains no $\star$'s. Second, $G$ encodes a function that maps the sizes of the domains in $\phi$ to the WMC of $\phi$ (more on this in \cref{sec:evaluation}).

\paragraph{TODO.}
\begin{itemize}
\item provide a short explanation of the types (emphasising which ones are new/updated). %\footnote{The new (as well as significantly renewed) types in \mathcal{T} are described in subsequent (TODO: which ones?) sections. The meanings of other types are described...}
\item have an example of a simple FCG
\end{itemize}

\subsection{Everything Else}

\begin{definition}
  A \emph{state} (of the search for an FCG for a given formula) is a tuple $(G, C, L)$, where:
  \begin{itemize}
  \item $G$ is an FCG (or \texttt{null}),
  \item $C$ is a compilation cache that maps integers to sets of pairs $(\phi, v)$, where $\phi$ is a formula, and $v$ is a vertex of $G$ (which is used to identify opportunities for recursion),
  \item and $L$ is a list of formulas (that are yet to be compiled). (Note that the order is crucial!)
  \end{itemize}
\end{definition}

\begin{definition}
  A (compilation) \emph{rule} is a function that takes a formula and returns a set of $(G, L)$ pairs, where $G$ is a (potentially \texttt{null}) FCG, and $L$ is a list of formulas. TODO: add an example showing that it's usually an FCG with one vertex and a bunch of $\star$'s marking a fixed number of outgoing edges.
\end{definition}

We assume that if there is a pair $(\texttt{null}, L)$ in the set returned by a rule, then $|L| = 1$, i.e., the rule transformed the formula without creating any vertices.

\begin{algorithm}[t]
  \caption{The (main part of the) search algorithm}
  \label{alg:search}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwData{solutions}{solutions}
  \SetKwFunction{applyGreedyRules}{applyGreedyRules}
  \SetKwFunction{applyAllRules}{applyAllRules}
  \SetKwFunction{put}{put}
  \SetKwFunction{get}{get}
  \SetKwFunction{emptyy}{empty}
  \KwIn{a formula $\phi_0$}
  \KwOut{all found FCGs for $\phi_0$ are in the set \solutions}
  $\solutions \gets \emptyset$\;
  $C_0 \gets \emptyset$\;
  $(G_0, C_0, L_0) \gets \applyGreedyRules{$\phi_0$, $C_0$}$\;
  \lIf{$L_0 = []$}{$\solutions \gets \{\, G_0 \,\}$}
  \Else{
    $q \gets \text{an empty queue of states}$\;
    $q.\put{$(G_0, C_0, L_0)$}$\;
    \While{{\bf not} $q.\emptyy{}$}{
      \ForEach{state $(G, C, L) \in \applyAllRules{$q.\get{}$}$}{
        \lIf{$L = []$}{$\solutions \gets \solutions \cup \{\, G \,\}$}
        \lElse{$q.\put{$(G, C, L)$}$}
      }
    }
  }
\end{algorithm}

\begin{algorithm}
  \caption{Functions used in \cref{alg:search} for applying compilation rules}
  \label{alg:apply}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwData{newStates}{newStates}
  \SetKwFunction{applyGreedyRules}{applyGreedyRules}
  \SetKwFunction{applyAllRules}{applyAllRules}
  \SetKwFunction{applyGreedyRulesToAllFormulas}{applyGreedyRulesToAllFormulas}
  \SetKwFunction{mergeFcgs}{mergeFcgs}
  \SetKwFunction{updateCache}{updateCache}
  \KwData{a set of greedy rules $\Gamma$}
  \KwData{a set of non-greedy rules $\Delta$}
  \Fn{\applyGreedyRules{$\phi$, $C$}}{
    \ForEach{rule $r \in \Gamma$}{
      \If{$r(\phi) \ne \emptyset$}{
        $(G, L) \gets \text{any element of } r(\phi)$\;
        \lIf{$G = {\normalfont \texttt{null}}$}{\Return{\applyGreedyRules{the only element of $L$, $C$}}}
        \Else{
          $(V, s, N^+, \tau) \gets G$\;
          $C \gets \updateCache{$C$, $\phi$, $G$}$\;
          \Return{\applyGreedyRulesToAllFormulas{$G$, $C$, $L$}}\;
        }
      }
    }
    \Return{$({\normalfont \texttt{null}}, C, [\phi])$}\;
  }
  \Fn{\applyGreedyRulesToAllFormulas{$(V, s, N^+, \tau)$, $C$, $L$}}{
    \lIf{$L = []$}{\Return{$((V, s, N^+, \tau), C, L)$}}
    $N^+(s) \gets []$\;
    $L' \gets []$\;
    \ForEach{formula $\phi \in L$}{
      $(G', C, L'') \gets \applyGreedyRules{$\phi$, $C$}$\;
      $L' \gets L' \mdoubleplus L''$\;
      \lIf{$G' = {\normalfont \texttt{null}}$}{$N^+(s) \gets N^+(s) \mdoubleplus [\star]$}
      \Else{
        $(V', s', N', \tau') \gets G'$\;
        $V \gets V \sqcup V'$\;
        $N^+ \gets N^+ \sqcup N'$\;
        $N^+(s) \gets N^+(s) \mdoubleplus [s']$\;
        $\tau \gets \tau \sqcup \tau'$\;
      }
    }
    \Return{$((V, s, N^+, \tau), C, L')$}\;
  }
  \Fn{\applyAllRules{$s$}}{
    $(G, C, L) \gets s$\;
    $\phi : T \gets L$\;
    $(G', C', L') \gets \text{a copy of } s$\;
    $\newStates \gets []$\;
    \ForEach{rule $r \in \Delta$}{
      \ForEach{$(G'', L'') \in r(\phi)$}{
        \lIf{$G'' = {\normalfont \texttt{null}}$}{$\newStates \gets \newStates \mdoubleplus \applyAllRules{$(G', C', L'')$}$}
        \Else{
          $(V, s, N^+, \tau) \gets G''$\;
          $C' \gets \updateCache{$C'$, $\phi$, $G''$}$\;
          $(G'', C', L'') \gets \applyGreedyRulesToAllFormulas{$G''$, $C'$, $L''$}$\;
          \lIf{$G' = {\normalfont \texttt{null}}$}{$\newStates \gets \newStates \mdoubleplus [(G'', C', L'' \mdoubleplus T)]$}
          \lElse{$\newStates \gets \newStates \mdoubleplus [(\mergeFcgs{$G'$, $G''$}, C', L'' \mdoubleplus T)]$}
        }
      }
      $(G', C', L') \gets \text{a copy of } s$\;
    }
    \Return{\newStates}\;
  }
\end{algorithm}

\begin{algorithm}
  \caption{Helper functions used by \cref{alg:apply}}
  \label{alg:helpers}
  \SetKwProg{Fn}{Function}{:}{}
  \SetKwFunction{mergeFcgs}{mergeFcgs}
  \SetKwFunction{updateCache}{updateCache}
  \Fn{\updateCache{$C$, $\phi$, $(V, s, N^+, \tau)$}}{
    \lIf{$\tau(s) = \textsc{Ref}$}{\Return{$C$}}
    \lIf{$\#\phi \not\in \dom(C)$}{\Return{$C \cup \{\, \#\phi \mapsto (\phi, s) \,\}$}}
    \lIf{there is no $(\phi', v) \in C(\#\phi)$ s.t. $v = s$}{$C(\#\phi) \gets (\phi, s) \mdoubleplus C(\#\phi)$}
    \Return{$C$}\;
  }
  \Fn{\mergeFcgs{$G = (V, s, N^+, \tau)$, $G' = (V', s', N', \tau')$, $r = s$}}{
    \lIf{$\tau(r) = \textsc{Ref}$}{\Return{\normalfont \texttt{null}}}
    \ForEach{$t \in N^+(r)$}{
      \If{$t = \star$}{
        replace $t$ with $s'$ in $N^+(r)$\;
        \Return{$(V \sqcup V', s, N^+ \sqcup N', \tau \sqcup \tau')$}\;
      }
      $G'' \gets \mergeFcgs{$G$, $G'$, $t$}$\;
      \lIf{$G'' \ne {\normalfont \texttt{null}}$}{\Return{$G''$}}
    }
    \Return{\normalfont \texttt{null}}\;
  }
\end{algorithm}

TODO: explain the `tail' part of the algorithm, i.e., that the first formula is replaced by some vertices and some formulas. And explain why we don't want to have \textsc{Ref} vertices in the cache.

Note: At the end, \texttt{mergeFcgs} will never return \texttt{null} because there is going to be at least one $\star$ in $G$ and the function will find it.

\section{Smoothing}

\begin{algorithm}
  \SetKwData{changed}{changed}
  \caption{Propagating atoms for smoothing across the FCG in a way that avoids infinite loops}
  \label{alg:smoothing}
  \KwIn{FCG $(V, s, N^+, \tau)$}
  \KwIn{function $\iota$ that maps vertex types in $\mathcal{T}$ to sets of atoms}
  \KwIn{functions $\{\,f_t\,\}_{t \in \mathcal{T}}$ that take a list of sets of atoms and return a set of atoms}
  \KwOut{function $S$ that maps vertices in $V$ to sets of atoms}
  $S \gets \{\, v \mapsto \iota(\tau(v)) \mid v \in V \,\}$\;
  $\changed \gets \texttt{true}$\;
  \While{\changed}{
    $\changed \gets \texttt{false}$\;
    \ForEach{vertex $v \in V$}{
      $S' \gets f_{\tau(v)}([S(w) \mid w \in N^+(v)])$\;
      \If{$S' \ne S(v)$}{
        $\changed \gets \texttt{true}$\;
        $S(v) \gets S'$\;
      }
    }
  }
\end{algorithm}

[Insert motivation for smoothing from Section 3.4. of the ForcLift paper.] Originally, smoothing was (and still is) a two-step process. First, atoms that are still accounted for in the circuit are propagated upwards. Then, at vertices of certain types, missing atoms are detected and additional sinks are created to account for them. If left unchanged, the first step of this process would result in an infinite loop whenever a cycle is encountered. \Cref{alg:smoothing} outlines how the first step can be adapted to an arbitrary directed graph.

\section{Identifying Possibilities for Recursion}

\paragraph{Notation.}
For any clause $c = (L, C, \delta_c)$, bijection $\beta\colon \Vars(c) \twoheadrightarrowtail V$ (for some set of variables $V$) and function $\gamma\colon \Doms(c) \to \mathcal{D}$, let $c[\beta, \gamma] = d$ be the clause $c$ with all occurrences of any variable $v \in \Vars(c)$ in $L$ and $C$ replaced with $\beta(v)$ (so $\Vars(d) = V$) and $\delta_d\colon V \to \mathcal{D}$ defined as $\delta_d \coloneqq \gamma \circ \delta_c \circ \beta^{-1}$. In other words, $\delta_d$ is the unique function that makes
\[
\begin{tikzcd}
  \Vars(c) \ar[r, tail, two heads, "\beta"] \arrow[d, swap, "\delta_c"] & V = \Vars(d) \ar[d, dashed, "\exists!\delta_d"] \\
  \Doms(c) \ar[r, swap, "\gamma"] & \mathcal{D}
\end{tikzcd}
\]
commute.

\begin{algorithm}[t]
  \caption{The compilation rule for $\Reff$}
  \label{alg:trycache}

  \SetKwProg{Fn}{Function}{:}{}
  \SetKw{KwRet}{yield}
  \SetKwData{compilationCache}{compilationCache}
  \SetKwFunction{identifyRecursion}{identifyRecursion}
  \SetKwFunction{generateMaps}{generateMaps}
  \SetKwFunction{constructDomainMap}{constructDomainMap}

  \KwIn{formula $\phi$}
  \KwOut{either a singleton with the new $\Reff$ vertex or $\emptyset$}

  \ForEach{$(\psi, v) \in \compilationCache{$\#\phi$}$}{
    $\rho \gets \identifyRecursion{$\phi$, $\psi$}$\;
    \lIf{$\rho \ne {\normalfont \texttt{null}}$}{\Return{$\{\, (\Reff_\rho(v), \emptyset) \,\}$}}
  }
  \Return{$\emptyset$}\;

  \Fn{\identifyRecursion{$\phi$, $\psi$, $\rho = \emptyset$}}{
    \lIf{$|\phi| \ne |\psi|$ {\bf or} $\#\phi \ne \#\psi$}{\Return{\normalfont \texttt{null}}}
    \lIf{$\phi = \emptyset$}{\Return{$\rho$}}
    \ForEach{clause $c \in \phi$\label{line:for1}}{
      \ForEach{clause $d \in \psi$ s.t. $\#d = \#c$\label{line:for2}} {
        \ForEach{$(\beta, \gamma) \in \generateMaps{$c$, $d$, $\rho$}$ s.t. $c[\beta, \gamma] = d$\label{line:generateMaps}}{
          $\rho' \gets \identifyRecursion{$\phi \setminus \{\, c \,\}$, $\psi \setminus \{\, d \,\}$, $\rho \cup \gamma$}$\;\label{line:recursion}
          \lIf{$\rho' \ne {\normalfont \texttt{null}}$}{\Return{$\rho'$}}
        }
      }
      \Return{\normalfont \texttt{null}}\;
    }
  }

  \Fn{\generateMaps{$c$, $d$, $\rho$}}{
    \ForEach{bijection $\beta\colon \Vars(c) \to \Vars(d)$\label{line:bijection}}{
      $\gamma \gets \constructDomainMap{$\Vars(c)$, $\delta_c$, $\delta_d$, $\beta$, $\rho$}$\;
      \lIf{$\gamma \ne {\normalfont \texttt{null}}$}{\KwRet{$(\beta, \gamma)$}}
    }
  }

  \Fn{\constructDomainMap{$V$, $\delta_c$, $\delta_d$, $\beta$, $\rho$}}{
    $\gamma \gets \emptyset$\;
    \ForEach{$v \in V$}{
      \lIf{$\delta_c(v) \in \dom(\rho)$ {\bf and} $\rho(\delta_c(v)) \ne \delta_d(\beta(v))$}{\Return{\normalfont \texttt{null}}}
      \lIf{$\delta_c(v) \not\in \dom(\gamma)$}{$\gamma \gets \gamma \cup \{\, \delta_c(v) \mapsto \delta_d(\beta(v)) \,\}$}
      \lElseIf{$\gamma(\delta_c(v)) \ne \delta_d(\beta(v))$}{\Return{\normalfont \texttt{null}}}
    }
    \Return{$\gamma$}\;
  }
\end{algorithm}

The function \texttt{traceAncestors} returns \texttt{null} if domain $d \in \mathcal{D}$ is not an ancestor of domain $e \in \mathcal{D}$. Otherwise, it returns \texttt{true} if the size of $e$ is guaranteed to be strictly smaller than the size of $d$ (i.e., there is domain created by the constraint removal rule on the path from $d$ to $e$) and \texttt{false} if their sizes will be equal at some point during inference.

Notation: For partial functions $\alpha, \beta\colon A \pfun B$ s.t. $\alpha|_{\dom(\alpha) \cap \dom(\beta)} = \beta|_{\dom(\alpha) \cap \dom(\beta)}$, we write $\alpha \cup \beta$ for the unique partial function s.t. $\alpha \cup \beta|_{\dom(\alpha)} = \alpha$, and $\alpha \cup \beta|_{\dom(\beta)} = \beta$. TODO: explain $\sqcup$ for both sets and functions.

\paragraph{TODO}
\begin{itemize}
\item update the example to not refer to things that don't exist anymore
\item introduce/describe and \cref{alg:trycache} and describe the cache that's being used.
\item explain why $\rho \cup \gamma$ is possible
\item explain what the second return statement is about and why a third one is not necessary
\item explain the yield keyword
\item in the example below: write down both formula using the ForcLift format
\end{itemize}

The algorithm could be improved in two ways:
\begin{itemize}
\item by constructing a domain map first and then using it to reduce the number of viable variable bijections.
\item by similarly using the domain map $\rho$.
\end{itemize}
However, it is not clear that this would result in an overall performance improvement, since the number of variables in instances of interest never exceeds three and the identity bijection is typically the right one.

Diagramatically, \texttt{constructDomainMap} attempts to find $\gamma\colon \Doms(c) \to \Doms(d)$ s.t. the following diagram commutes (and returns \texttt{null} if such a function does not exist):
\[
\begin{tikzcd}
  \Vars(c) \ar[r, tail, two heads, "\beta"] \arrow[d, swap, "\delta_c"] & \Vars(d) \ar[d, "\delta_d"] \\
  \Doms(c) \ar[r, dashed, "\gamma"] \ar[d, hookrightarrow] & \Doms(d) \ar[d, hookrightarrow] \\
  \mathcal{D} \ar[r, swap, "\rho"] & \mathcal{D}.
\end{tikzcd}
\]

\begin{example} \label{example}
  Let $\phi$ be the formula
  \begin{align}
    \forall X \in a. \forall Y \in b. \forall Z \in b. Y \ne Z &\implies \neg p(X, Y) \lor \neg p(X, Z)\label{phi1} \\
    \forall X \in a. \forall Y \in b. \forall Z \in a. X \ne Z &\implies \neg p(X, Y) \lor \neg p(Z, Y).\label{phi2}
  \end{align}
  and $\psi$ be the formula
  \begin{align}
    \forall X \in a'. \forall Y \in b^\bot. \forall Z \in b^\bot. Z \ne Y &\implies \neg p(X, Y) \lor \neg p(X, Z)\label{psi1} \\
    \forall X \in a'. \forall Y \in b^\bot. \forall Z \in a'. X \ne Z &\implies \neg p(X, Y) \lor \neg p(Z, Y)\label{psi2}
  \end{align}

  The relevant domains and the definition of $\pi$ is in \cref{fig:example}. Since $\#\phi = \#\psi$, and the formulas are non-empty, the algorithm proceeds with the for-loops on \cref{line:for1,line:for2,line:generateMaps}. Suppose $c$ in the algorithm refers to \cref{phi1}, and $d$ to \cref{psi1}. Since both clauses have three variables, in the worst case, function \texttt{generateMaps} would have $3!=6$ bijections to check. Suppose the identity bijection is picked first. Then \texttt{constructDomainMap} is called with the following parameters:
  \begin{itemize}
  \item $V = \{\, X, Y, Z \,\}$,
  \item $\delta_c = \{\, X \mapsto a, Y \mapsto b, Z \mapsto b \,\}$,
  \item $\delta_d = \{\, X \mapsto a', Y \mapsto b^\bot, Z \mapsto b^\bot \,\}$,
  \item $\beta = \{\, X \mapsto X, Y \mapsto Y, Z \mapsto Z \,\}$,
  \item $\rho = \emptyset$.
  \end{itemize}
  Since $\delta_i(Y) = \delta_i(Z)$ for $i \in \{\, c, d \,\}$, \texttt{constructDomainMap} returns $\gamma = \{\, a \mapsto a', b \mapsto b^\bot \,\}$. Thus, \texttt{generateMaps} yields its first pair of maps $(\beta, \gamma)$ to \cref{line:generateMaps}. Furthermore, the pair satisfies $c[\beta, \gamma] = d$.

  Since $\pi(a') = a$, and $a' \in \mathcal{C}$, \texttt{traceAncestors($a$, $a'$)} returns \texttt{true}, which sets $\textsf{foundConstraintRemoval}'$ to \texttt{true} as well. When $e = b$, however, \texttt{traceAncestors}($b$, $b^\bot$) returns \texttt{false} since $b^\bot$ is a descendant of $b$ but not created by the constraint removal compilation rule. On \cref{line:recursion}, a recursive call to \texttt{identifyRecursion($\phi'$, $\psi'$, $\gamma$, true)} is made, where $\phi'$ and $\psi'$ are new formulas with one clause each: \cref{phi2} and \cref{psi2}, respectively.

  Again we have two non-empty formulas with equal hash codes, so \texttt{generateMaps} is called with $c$ set to \cref{phi2}, $d$ set to \cref{psi2}, and $\rho = \{\, a \mapsto a', b \mapsto b^\bot \,\}$. Suppose \cref{line:bijection} picks the identity bijection first again. Then \texttt{constructDomainMap} is called with the following parameters:
  \begin{itemize}
  \item $V = \{\, X, Y, Z \,\}$,
  \item $\delta_c = \{\, X \mapsto a, Y \mapsto b, Z \mapsto a \,\}$,
  \item $\delta_d = \{\, X \mapsto a', Y \mapsto b^\bot, Z \mapsto a' \,\}$,
  \item $\beta = \{\, X \mapsto X, Y \mapsto Y, Z \mapsto Z \,\}$,
  \item $\rho = \{\, a \mapsto a', b \mapsto b^\bot \,\}$.
  \end{itemize}
  Since $\beta$ and $\rho$ `commute' (TODO: as in the diagram above), and there are no new domains in $\Doms(c)$ and $\Doms(d)$, $\gamma$ exists and is equal to $\rho$. Again, the returned pair $(\beta, \gamma)$ satisfies $c[\beta, \gamma] = d$. This $\gamma$ passes the \texttt{traceAncestors} checks exactly the same way as the one before, and \cref{line:recursion} calls \texttt{identifyRecursion($\emptyset$, $\emptyset$, $\rho$, true)}, which immediately returns $\rho = \{\, a \mapsto a', b \mapsto b^\bot \,\}$ as the final answer. This means that one can indeed use an FCG for $\WMC(\phi)$ to compute $\WMC(\psi)$ by replacing every mention of $a$ with $a'$ and every mention of $b$ with $b^\bot$.
\end{example}

\subsection{Evaluation}

$\WMC(\Reff_\rho(v); \size) = \WMC(v; \size')$ ($n$ is the target vertex), where $\size'$ is defined as
\[
\size'(x) =
\begin{cases}
  \size(\rho(x)) & \text{if } x \in \dom(\rho) \\
  \size(x) & \text{otherwise}
\end{cases}
\]
for all $x \in \mathcal{D}$.

\section{New Compilation Rules}

Throughout this section, let $\phi$ be an arbitrary formula.

\subsection{Constraint Removal}

\begin{algorithm}[t]
  \caption{The compilation rule for $\CR$}
  \label{alg:constraintremoval}
  \KwIn{formula $\phi$, set of domains $\mathcal{D}$}
  \KwOut{set $S$}
  $S \gets \emptyset$\;
  \ForEach{domain $d \in \mathcal{D}$ and element $e \in d$ s.t. $e$ does not occur in any literal of any clause of $\phi$ {\bf and} for each clause $ c = (L, C, \delta_c) \in \phi$ and variable $v \in \Vars(c)$, either $\delta_c(v) \ne d$ {\bf or} $(v, e) \in C$}{
    add a new domain $d'$ to $\mathcal{D}$\;
    $\phi' \gets \emptyset$\;
    \ForEach{clause $(L, C, \delta) \in \phi$}{
      $C' \gets \{\, (x, y) \in C \mid y \ne e \,\}$\;
      $\delta' \gets \emptyset$\;
      \ForEach{variable $v \in \Vars(L) \cup \Vars(C')$}{
        \lIf{$\delta(v) = d$}{$\delta' \gets \delta' \cup \{\, v \mapsto d' \,\}$}
        \lElse{$\delta' \gets \delta' \cup \{\, v \mapsto \delta(v) \,\}$}
      }
      $\phi' \gets \phi' \cup \{\, (L, C', \delta') \,\}$\;
    }
    $S \gets S \cup \{\, \CR_{d \mapsto d'}, \phi' \,\}$\;
  }
\end{algorithm}

TODO: describe \cref{alg:constraintremoval} and rewrite the example below.

\begin{example}
  Let $\phi = \{\, c_1, c_2, c_e \,\}$ be a formula with clauses (constants lowercase, variables uppercase)
  \begin{align*}
    c_1 &= (\emptyset, \{\, (Y, X) \,\}, \{\, X \mapsto b^\top, Y \mapsto b^\top \,\}), \\
    c_2 &= (\{\, \neg p(X, Y), \neg p(X, Z) \,\}, \{\, (X, x), (Y, Z) \,\}, \{\, X \mapsto a, Y \mapsto b^\bot, Z \mapsto b^\bot \,\}), \\
    c_3 &= (\{\, \neg p(X, Y), \neg p(Z, Y) \,\}, \{\, (X, x), (Z, X), (Z, x) \,\}, \{\, X \mapsto a, Y \mapsto b^\bot, Z \mapsto a \,\}).
  \end{align*}
  Domain $a$ and with its element $x \in a$ satisfy the preconditions for constraint removal. The operator introduces a new domain $a'$ and transforms $\phi$ to $\phi' = (c_1', c_2', c_3')$, where
  \begin{align*}
    c_1' &= c_1 \\
    c_2' &= (\{\, \neg p(X, Y), \neg p(X, Z) \,\}, \{\, (Y, Z) \,\}, \{\, X \mapsto a', Y \mapsto b^\bot, Z \mapsto b^\bot \,\}) \\
    c_3' &= (\{\, \neg p(X, Y), \neg p(Z, Y) \,\}, \{\, (Z, X) \,\}, \{\, X \mapsto a', Y \mapsto b^\bot, Z \mapsto a' \,\}).
  \end{align*}
\end{example}

\paragraph{Evaluation.}
\[
\WMC(\CR_{d \mapsto d'}(n); \size) =
\begin{cases}
  \WMC(n; \size \cup \{\, d' \mapsto \size(d) - 1 \,\}) & \text{if } \size(d) > 0 \\
  0 & \text{otherwise.}
\end{cases}
\]

\subsection{A Generalisation of Domain Recursion}

\begin{algorithm}[t]
  \caption{The compilation rule for $\DR$}
  \label{alg:domainrecursion}
  \KwIn{formula $\phi$}
  \KwOut{set $S$}
  $S \gets \emptyset$\;
  \ForEach{domain $d \in \mathcal{D}$ s.t. there is a clause $c \in \phi$ and a variable $v \in \Vars(L_c)$ s.t. $\delta_c(v) = d$}{
    $\phi' \gets \emptyset$\;
    $x \gets \text{a new constant associated with domain } d$\;
    \ForEach{clause $c = (L, C, \delta) \in \phi$\label{line:forclause}}{
      $V \gets \{\, v \in \Vars(L) \mid \delta(v) = d \,\}$\;
      \ForEach{subset $W \subseteq V$ s.t. $W^2 \cap C = \emptyset$ {\bf and} $W \cap \{\, v \in \Vars(C) \mid (v, y) \in C \text{ for some constant } y \,\} = \emptyset$\label{line:conditions}}{
        \tcc{Here, $\delta'$ is the restriction of $\delta$ to the new set of variables}
        $\phi' \gets \phi' \cup \{\, (L[x/W], C[x/W] \cup \{\, (v, x) \mid (v \in V \setminus W) \,\}, \delta') \,\}$\;
      }
    }
    $S \gets S \cup \{\, (\DR_d, \phi') \,\}$\;
  }
\end{algorithm}

\paragraph{The algorithm uses this notation for substitution.}
Let $S$ be a set of constraints or literals, $V$ a set of variables, and $x$ either a variable or a constant. Then we write $S[x/V]$ to denote $S$ with all occurrences of all variables in $V$ replaced with $x$.\footnote{Note that if $(v, w)$ is a two-variable constraint, substituting a constant $c$ for $v$ would result in $(c, w)$, which would have to be rewritten as $(w, c)$ to fit the definition of a constraint.}

TODO: Compare with the original \cite{DBLP:conf/nips/Broeck11}.

The reason for this precondtion is the same as in the initial version of domain recursion: there must be a variable with that domain featured among the literals because it needs to be replaced by a constant. TODO: expand this.

TODO: describe \cref{alg:domainrecursion}.

\begin{example}
  Let $\phi = \{\, c_1, c_2 \,\}$ be a formula, where
  \begin{align*}
    c_1 &= (\{\, \neg p(X, Y), \neg p(X, Z) \,\}, \{\, (Z, Y) \,\}, \{ X \mapsto a, Y \mapsto b, Z \mapsto b \}), \\
    c_2 &= (\{\, \neg p(X, Y), \neg p(Z, Y) \,\}, \{\, (Z, X) \,\}, \{ X \mapsto a, Y \mapsto b, Z \mapsto a \}).
  \end{align*}
  While domain recursion is possible on both domains, here we illustrate how it works on $a$.

  Suppose \cref{line:forclause} picks $c = c_1$ first. Then $V = \{\, X \,\}$. Both subsets of $V$ satisfy the conditions on \cref{line:conditions} and generate new clauses
  \[
  (\{\, \neg p(X, Y), \neg p(X, Z) \,\}, \{\, (Z, Y), (X, x) \,\}, \{ X \mapsto a, Y \mapsto b, Z \mapsto b \}),
  \]
  (from $W = \emptyset$) and
  \[
  (\{\, \neg p(x, Y), \neg p(x, Z) \,\}, \{\, (Z, Y) \,\}, \{ Y \mapsto b, Z \mapsto b \})
  \]
  (from $W = V$).

  When $c = c_2$, then $V = \{\, X, Z \,\}$. The subset $W = V$ fails to satisfy the first condition because of the $Z \ne X$ constraint; without this condition, the resulting clause would have an unsatisfiable constraint $x \ne x$. The other three subsets of $V$ all generate clauses for $\phi'$:
  \[
  (\{\, \neg p(X, Y), \neg p(Z, Y) \,\}, \{\, (Z, X), (X, x), (Z, x) \,\}, \{ X \mapsto a, Y \mapsto b, Z \mapsto a \})
  \]
  (from $W = \emptyset$),
  \[
  (\{\, \neg p(x, Y), \neg p(Z, Y) \,\}, \{\, (Z, x) \,\}, \{ Y \mapsto b, Z \mapsto a \})
  \]
  (from $W = \{\, X \,\}$), and
  \[
  (\{\, \neg p(X, Y), \neg p(x, Y) \,\}, \{\, (X, x) \,\}, \{ X \mapsto a, Y \mapsto b, \})
  \]
  (from $W = \{\, Z \,\}$).
\end{example}

\paragraph{Evaluation.}
\[
\WMC(\DR_d(n); \size) =
\begin{cases}
  \WMC(n; \size) & \text{if } \size(d) > 0 \\
  1 & \text{otherwise.}
\end{cases}
\]

One is picked as the multiplicative identity.

\section{Other Topics}

\begin{itemize}
\item new rules that don't create vertices (e.g., duplicate removal, unconditional contradiction detection, etc.)
\item some notes on halting
  \begin{itemize}
  \item Search is infinite. Some rules increase the size of the formula(s), but most reduce it.
  \item Inference is guaranteed to terminate if at least one domain shrinks by at least one. But note that allowing recursive calls with the same domain sizes (e.g., $f(n) = f(n) + \dots$) could be useful because these problematic terms might cancel out.
  \item It's impossible for $n \gets n - 1$ and $\texttt{for } n \in \dots$ to combine in a way that results in an infinite loop.
  \end{itemize}
\item care should be taken when cloning to preserve the validity of the cache and avoid infinite cycles (we use a separate (node $\to$ node) cache for this)
\item another change: one rule can return more than one `continuation of the graph'
\end{itemize}

\section{How to Evaluate an FCG} \label{sec:evaluation}

Along with the three vertex types described above, here are all the other ones. This section is mostly just taken from \cite{DBLP:conf/ijcai/BroeckTMDR11} but with some changes in notation.

TODO: explain that $x, y, z$ refer to vertices, $c$ refers to a clause, and describe each vertex type in a bit more detail.

\begin{description}
\item[tautology] $\WMC(\Tautology; \size) = 1$
\item[contradiction] $\WMC(\Contradiction c; \size) = 0^{\gr(c; \size)}$
\item[unit clause]
  \[
  \WMC(\Unit c; \size) =
  \begin{cases}
    \wwp(p)^{\gr(c; \size)} & \text{if the literal is positive} \\
    \wwn(p)^{\gr(c; \size)} & \text{otherwise,}
  \end{cases}
  \]
  where $p$ is the predicate of the literal.
\item[smoothing] $\WMC(\Smoothing c; \size) = (\wwp(p) + \wwn(p))^{\gr(c; \size)}$, where $p$ is the predicate of the literal.
\item[decomposable conjunction] $\WMC(x \oland y; \size) = \WMC(x; \size) \times \WMC(y; \size)$
\item[deterministic disjunction] $\WMC(x \olor y; \size) = \WMC(x; \size) + \WMC(y; \size)$
\item[decomposable set-conjunction] $\WMC(\oland_D x; \size) = \WMC(x; \size)^{\size(D)}$
\item[deterministic set-disjunction] $\WMC(\olor_{D \subseteq S} x; \size) = \sum_{d = 0}^{\size(S)} \binom{\size(S)}{d} \WMC(x; \size \cup \{\, D \mapsto d \,\})$
\item[inclusion-exclusion] $\WMC(\IE(x, y, z); \size) = \WMC(x; \size) + \WMC(y; \size) - \WMC(z; \size)$
\end{description}

\section{Newly Domain-Liftable Formulas}

\begin{figure}
  \centering
  \begin{tikzpicture}[every node/.style={draw,ellipse},edge from parent/.style={draw,-latex},sibling distance=48mm]
    \node (dr) {$\DR_{x \in a}$}
    child {node {$\bigvee_{b^\top \subseteq b}$}
      child {node {$\wedge$}
        child {node {$1_{(\{\, p(x, X) \,\}, \emptyset, \{\, X \mapsto b^\top \,\})}$}}
        child {node {$\wedge$}
          child {node {$1_{(\{\, \neg p(x, X) \,\}, \emptyset, \{\, X \mapsto b^\bot \,\})}$}}
          child {node {$\wedge$}
            child {node {$1_{(\{\, \neg p(X, Y) \,\}, \{\, (X, x) \,\}, \{\, X \mapsto a, Y \mapsto b^\top \,\})}$}}
            child {node {$\CR_{a \mapsto a'}$}
              child {node {$\wedge$}
                child {node {$\bot_{(\emptyset, \{\, (X, Y) \,\}, \{\, X \mapsto b^\top, Y \mapsto b^\top \,\})}$}}
                child {node (ref) {$\Reff_{\{\, a \mapsto a', b \mapsto b^\bot \,\}}$}}
              }}}}}};
    \draw[-latex, bend right] (ref) to (dr);
  \end{tikzpicture}
  \caption{A graphical representation of an FCG for injections and partial injections between two domains. TODO: add a reference to a formula?}
\end{figure}

\paragraph{TODO.}
\begin{itemize}
\item $b^\bot$ should be part of the notation for counting.
\item \textsc{ForcLift} fails on all of these.
\item Mention that functions, surjections, and their partial counterparts are/were already liftable.
\item Formulas for partial injections, how to add totality, surjectivity.
\item Describe the figure and connect it to the algebraic formula.
\item Explain the algebraic notation that I'm using here (e.g., that $f$ is always the main function)
\item Explain the Knuth's bracket notation (or use something else, since I'm already using brackets for lists)
\item Explain the importance of comparing domain sizes to 2.
\item Maybe a big table: name, formula, optimal solution and its complexity, best found solution, its depth, its complexity. Also mention that it only takes a few seconds to find these solutions.
\item Going beyond depth 6 (or sometimes even completing depth 6) is computationally infeasible with the current implementation, but depth at most 5 can be searched within at most a few seconds.
\item Notation for domain recursion should include the constant name.
\item Combine the tikz and the algebraic notation into one, so I don't need to have two versions. But how? Maybe associate a symbol with each type and only to the types that I use?
\item Partial bijection is the same as partial injection.
\item the exponential solutions can be computed in quadratic time with dynamic programming!
\end{itemize}

\begin{remark}
  FCGs that compare the size of a domain to an integer can be constructed automatically using compilation rules, although $n$ is upper bounded by the maximum number of variables in any clause of the input formula since there is no rule that would introduce new variables.
\end{remark}

\begin{itemize}
\item 1d bijections and 1d injections (note that it's the same problem). Depth 3 solution:
  \begin{align*}
    f(n) &= \sum_{m=0}^n \binom{n}{m} (-1)^{n-m}g(n, m) \\
    g(n, m) &= \sum_{l=0}^n \binom{n}{l}[l < 2]g(n-l, m-1) \\
    &= g(n, m - 1) + ng(n - 1, m - 1),
  \end{align*}
  which works with base case $g(n, 0) = 1$.
\item 1d partial injections. 2 solutions at depth 6, but they're too complicated to check by hand. A contradiction with $X \ne x$ constraints makes things complicated.
\item 2d bijections. Depth 3:
  \begin{align*}
    f(m, n) &= \sum_{l=0}^m \binom{m}{l} [l < 2] (1 - [l < 1])f(m-l, n-1) \\
    &= mf(m-1, n-1),
  \end{align*}
  which works with base cases $f(0, 0) = 1$, $f(0, n) = 0$, $f(m, 0) = 0$.
\item 2d injections. Depth 2:
  \begin{align*}
    f(m, n) &= \sum_{l=0}^m \binom{m}{l}[l<2]f(m-l, n-1) \\
    &= f(m, n-1) + mf(m-1, n-1),
  \end{align*}
  which works with base cases $f(0, 0) = 1$ and $f(m, 0) = 0$.
\item 2d partial injections, depth 2. Exactly the same circuit as above but with base case $f(m, 0) = 1$.
\end{itemize}

\section{Conclusions and Future Work}

\paragraph{Conclusions and observations.}
\begin{itemize}
\item $\CR$ must be separate from $\DR$ because initially the requirement to not have the newly introduced constant in the literals is not satisfied.
\end{itemize}

\paragraph{Future work.}
\begin{itemize}
\item Transform FCGs to definitions of (possibly recursive) functions on integers. Use a computer algebra system to simplify them.
\item Design an algorithm to infer the necessary base cases. (Note that there can be an infinite amount of them when functions have more than one parameter.)
\item Observation: -1 (and powers thereof) appear in every solution to a formula if and only if the formula has existential quantification. That's not very smart! By putting unit propagation into $\Gamma$, these powers are pushed to the outer layers of the solution (i.e., `early' in the FCG). It's likely that removing this restriction would enable the algorithm to find asymptotically optimal solutions.
\end{itemize}

\bibliographystyle{acm}
\bibliography{notes}

\end{document}