### 2023-03-08 doc/paper/plot.tex
An experiment

% Created by tikzDevice version 0.12.3.1 on 2023-03-08 13:57:11
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (239.21,148.15);
\begin{scope}
\path[clip] ( 36.11, 30.69) rectangle (233.71,142.65);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11, 51.57) --
	(233.71, 51.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11, 81.09) --
	(233.71, 81.09);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11,110.61) --
	(233.71,110.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11,140.13) --
	(233.71,140.13);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 59.76, 30.69) --
	( 59.76,142.65);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 96.42, 30.69) --
	( 96.42,142.65);

\path[draw=drawColor,line width= 0.3pt,line join=round] (133.08, 30.69) --
	(133.08,142.65);

\path[draw=drawColor,line width= 0.3pt,line join=round] (169.74, 30.69) --
	(169.74,142.65);

\path[draw=drawColor,line width= 0.3pt,line join=round] (206.40, 30.69) --
	(206.40,142.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11, 36.81) --
	(233.71, 36.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11, 66.33) --
	(233.71, 66.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11, 95.85) --
	(233.71, 95.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11,125.37) --
	(233.71,125.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.43, 30.69) --
	( 41.43,142.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.09, 30.69) --
	( 78.09,142.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.75, 30.69) --
	(114.75,142.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.41, 30.69) --
	(151.41,142.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (188.07, 30.69) --
	(188.07,142.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (224.73, 30.69) --
	(224.73,142.65);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 45.09, 36.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 48.76, 36.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 52.43, 36.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 56.09, 36.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 59.76, 36.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 63.42, 36.82) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 67.09, 36.82) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 70.76, 36.82) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 74.42, 36.83) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 78.09, 36.83) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 81.75, 36.84) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 85.42, 36.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 89.09, 36.87) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 92.75, 36.89) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 96.42, 36.92) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (100.08, 37.06) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (103.75, 37.00) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (107.42, 37.05) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (111.08, 37.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.75, 37.20) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (118.42, 37.30) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (122.08, 37.44) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (125.75, 37.58) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (129.41, 37.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (133.08, 38.01) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (136.75, 38.27) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (140.41, 38.63) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (144.08, 39.00) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (147.74, 39.52) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (151.41, 40.24) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (155.08, 40.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (158.74, 41.68) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (162.41, 43.30) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (166.07, 43.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (169.74, 45.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (173.41, 47.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (177.07, 48.94) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (180.74, 51.21) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (184.40, 53.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (188.07, 57.94) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (191.74, 62.58) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (195.40, 65.03) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (199.07, 71.20) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (202.74, 75.90) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (206.40, 82.86) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (210.07, 93.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (213.73,100.82) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (217.40,111.86) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (221.07,123.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (224.73,137.56) circle (  1.96);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 45.09, 38.60) --
	( 47.37, 37.94) --
	( 49.64, 37.39) --
	( 51.91, 36.93) --
	( 54.19, 36.56) --
	( 56.46, 36.27) --
	( 58.74, 36.06) --
	( 61.01, 35.91) --
	( 63.28, 35.82) --
	( 65.56, 35.78) --
	( 67.83, 35.78) --
	( 70.11, 35.82) --
	( 72.38, 35.89) --
	( 74.65, 35.99) --
	( 76.93, 36.11) --
	( 79.20, 36.25) --
	( 81.48, 36.40) --
	( 83.75, 36.56) --
	( 86.02, 36.72) --
	( 88.30, 36.88) --
	( 90.57, 37.04) --
	( 92.85, 37.20) --
	( 95.12, 37.35) --
	( 97.39, 37.49) --
	( 99.67, 37.62) --
	(101.94, 37.74) --
	(104.21, 37.84) --
	(106.49, 37.94) --
	(108.76, 38.02) --
	(111.04, 38.08) --
	(113.31, 38.14) --
	(115.58, 38.18) --
	(117.86, 38.22) --
	(120.13, 38.24) --
	(122.41, 38.26) --
	(124.68, 38.27) --
	(126.95, 38.29) --
	(129.23, 38.30) --
	(131.50, 38.32) --
	(133.78, 38.35) --
	(136.05, 38.39) --
	(138.32, 38.45) --
	(140.60, 38.53) --
	(142.87, 38.64) --
	(145.15, 38.79) --
	(147.42, 38.97) --
	(149.69, 39.20) --
	(151.97, 39.48) --
	(154.24, 39.82) --
	(156.51, 40.23) --
	(158.79, 40.71) --
	(161.06, 41.27) --
	(163.34, 41.93) --
	(165.61, 42.69) --
	(167.88, 43.55) --
	(170.16, 44.53) --
	(172.43, 45.65) --
	(174.71, 46.90) --
	(176.98, 48.30) --
	(179.25, 49.86) --
	(181.53, 51.59) --
	(183.80, 53.51) --
	(186.08, 55.62) --
	(188.35, 57.94) --
	(190.62, 60.48) --
	(192.90, 63.25) --
	(195.17, 66.27) --
	(197.44, 69.56) --
	(199.72, 73.11) --
	(201.99, 76.96) --
	(204.27, 81.11) --
	(206.54, 85.59) --
	(208.81, 90.40) --
	(211.09, 95.56) --
	(213.36,101.08) --
	(215.64,107.00) --
	(217.91,113.31) --
	(220.18,120.05) --
	(222.46,127.22) --
	(224.73,134.85);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 45.09, 36.59) --
	( 47.37, 36.61) --
	( 49.64, 36.64) --
	( 51.91, 36.66) --
	( 54.19, 36.68) --
	( 56.46, 36.71) --
	( 58.74, 36.73) --
	( 61.01, 36.75) --
	( 63.28, 36.78) --
	( 65.56, 36.80) --
	( 67.83, 36.82) --
	( 70.11, 36.85) --
	( 72.38, 36.87) --
	( 74.65, 36.89) --
	( 76.93, 36.92) --
	( 79.20, 36.94) --
	( 81.48, 36.97) --
	( 83.75, 36.99) --
	( 86.02, 37.01) --
	( 88.30, 37.04) --
	( 90.57, 37.06) --
	( 92.85, 37.08) --
	( 95.12, 37.11) --
	( 97.39, 37.13) --
	( 99.67, 37.15) --
	(101.94, 37.18) --
	(104.21, 37.20) --
	(106.49, 37.22) --
	(108.76, 37.25) --
	(111.04, 37.27) --
	(113.31, 37.30) --
	(115.58, 37.32) --
	(117.86, 37.34) --
	(120.13, 37.37) --
	(122.41, 37.39) --
	(124.68, 37.41) --
	(126.95, 37.44) --
	(129.23, 37.46) --
	(131.50, 37.48) --
	(133.78, 37.51) --
	(136.05, 37.53) --
	(138.32, 37.55) --
	(140.60, 37.58) --
	(142.87, 37.60) --
	(145.15, 37.63) --
	(147.42, 37.65) --
	(149.69, 37.67) --
	(151.97, 37.70) --
	(154.24, 37.72) --
	(156.51, 37.74) --
	(158.79, 37.77) --
	(161.06, 37.79) --
	(163.34, 37.81) --
	(165.61, 37.84) --
	(167.88, 37.86) --
	(170.16, 37.88) --
	(172.43, 37.91) --
	(174.71, 37.93) --
	(176.98, 37.96) --
	(179.25, 37.98) --
	(181.53, 38.00) --
	(183.80, 38.03) --
	(186.08, 38.05) --
	(188.35, 38.07) --
	(190.62, 38.10) --
	(192.90, 38.12) --
	(195.17, 38.14) --
	(197.44, 38.17) --
	(199.72, 38.19) --
	(201.99, 38.21) --
	(204.27, 38.24) --
	(206.54, 38.26) --
	(208.81, 38.28) --
	(211.09, 38.31) --
	(213.36, 38.33) --
	(215.64, 38.36) --
	(217.91, 38.38) --
	(220.18, 38.40) --
	(222.46, 38.43) --
	(224.73, 38.45);
\end{scope}
\begin{scope}
\path[clip] ( 36.11, 30.69) rectangle (233.71,142.65);
\definecolor{drawColor}{RGB}{217,95,2}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round,fill=fillColor] (179.66,120.51) --
	(218.89,120.51) --
	(218.82,120.52) --
	(219.11,120.53) --
	(219.39,120.59) --
	(219.66,120.69) --
	(219.92,120.83) --
	(220.14,121.02) --
	(220.33,121.24) --
	(220.49,121.48) --
	(220.60,121.75) --
	(220.67,122.03) --
	(220.70,122.32) --
	(220.70,122.32) --
	(220.70,132.33) --
	(220.70,132.33) --
	(220.67,132.62) --
	(220.60,132.91) --
	(220.49,133.17) --
	(220.33,133.42) --
	(220.14,133.64) --
	(219.92,133.82) --
	(219.66,133.97) --
	(219.39,134.07) --
	(219.11,134.13) --
	(218.89,134.14) --
	(179.66,134.14) --
	(179.88,134.13) --
	(179.59,134.14) --
	(179.30,134.10) --
	(179.02,134.02) --
	(178.75,133.90) --
	(178.52,133.73) --
	(178.31,133.53) --
	(178.13,133.30) --
	(178.00,133.04) --
	(177.90,132.77) --
	(177.86,132.48) --
	(177.85,132.33) --
	(177.85,122.32) --
	(177.86,122.47) --
	(177.86,122.18) --
	(177.90,121.89) --
	(178.00,121.61) --
	(178.13,121.36) --
	(178.31,121.12) --
	(178.52,120.92) --
	(178.75,120.76) --
	(179.02,120.63) --
	(179.30,120.55) --
	(179.59,120.52) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] ( 36.11, 30.69) rectangle (233.71,142.65);
\definecolor{drawColor}{RGB}{217,95,2}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (199.27,123.53) {Quartic};
\definecolor{drawColor}{RGB}{27,158,119}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.3pt,line join=round,line cap=round,fill=fillColor] (185.61, 41.17) --
	(218.87, 41.17) --
	(218.79, 41.17) --
	(219.08, 41.18) --
	(219.37, 41.24) --
	(219.64, 41.34) --
	(219.89, 41.49) --
	(220.12, 41.67) --
	(220.31, 41.89) --
	(220.47, 42.14) --
	(220.58, 42.40) --
	(220.65, 42.69) --
	(220.67, 42.98) --
	(220.67, 42.98) --
	(220.67, 52.99) --
	(220.67, 52.99) --
	(220.65, 53.28) --
	(220.58, 53.56) --
	(220.47, 53.83) --
	(220.31, 54.07) --
	(220.12, 54.29) --
	(219.89, 54.48) --
	(219.64, 54.62) --
	(219.37, 54.72) --
	(219.08, 54.78) --
	(218.87, 54.80) --
	(185.61, 54.80) --
	(185.83, 54.78) --
	(185.54, 54.79) --
	(185.25, 54.76) --
	(184.97, 54.68) --
	(184.71, 54.55) --
	(184.47, 54.39) --
	(184.26, 54.19) --
	(184.09, 53.95) --
	(183.95, 53.70) --
	(183.86, 53.42) --
	(183.81, 53.13) --
	(183.81, 52.99) --
	(183.81, 42.98) --
	(183.81, 43.12) --
	(183.81, 42.83) --
	(183.86, 42.54) --
	(183.95, 42.27) --
	(184.09, 42.01) --
	(184.26, 41.78) --
	(184.47, 41.58) --
	(184.71, 41.41) --
	(184.97, 41.29) --
	(185.25, 41.21) --
	(185.54, 41.17) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] ( 36.11, 30.69) rectangle (233.71,142.65);
\definecolor{drawColor}{RGB}{27,158,119}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (202.24, 44.18) {Linear};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (239.21,148.15);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16, 33.78) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16, 63.30) {250};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16, 92.82) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16,122.34) {750};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (239.21,148.15);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 41.43, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 78.09, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (114.75, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (151.41, 19.68) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (188.07, 19.68) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (224.73, 19.68) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (239.21,148.15);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (134.91,  7.64) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (239.21,148.15);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08, 86.67) {Runtime (s)};
\end{scope}
\end{tikzpicture}

\end{document}