### 2023-03-17 doc/talks/3_long/talk.tex
Ditto

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{complexity}
\usepackage[style=authoryear]{biblatex}
\usepackage{forest}
\usepackage{soul}
\usepackage{csquotes}
\usepackage{pifont}

\usetikzlibrary{shapes}
\usetikzlibrary{positioning}
\usetikzlibrary{overlay-beamer-styles}
\usetikzlibrary{shapes.callouts}
\usetikzlibrary{cd}
\usetikzlibrary{calc}
\usetikzlibrary{backgrounds}
\usetikzlibrary{fit}

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\renewcommand<>{\hl}[1]{\only#2{\beameroriginal{\hl}}{#1}}
\makeatletter
\newcommand\SoulColor{%
  \let\set@color\beamerorig@set@color
  \let\reset@color\beamerorig@reset@color}
\makeatother
\SoulColor

\beamertemplatenavigationsymbolsempty
\usetheme{metropolis}
\setbeamercolor{background canvas}{bg=white}

\addbibresource{talk.bib}

\definecolor{color1}{HTML}{1B9E77}
\definecolor{color2}{HTML}{7570b3}
\definecolor{color3}{HTML}{a6761d}
\definecolor{forestgreen}{HTML}{009B55}

\DeclareMathOperator{\GDR}{\textsc{GDR}}
\DeclareMathOperator{\CR}{\textsc{CR}}
\DeclareMathOperator{\Reff}{\textsc{Ref}}

\colorlet{redish}{color2!25}
\sethlcolor{redish}
\setstcolor{black}

\tikzset{
  inner/.style={
    ellipse,
    draw,
  },
  inner on/.style={alt=#1{inner}{}},
  leaf/.style={
    rectangle,
    draw,
    fill=gray!25,
  },
  leaf on/.style={alt=#1{leaf}{}},
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt=#1{}{invisible}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
  },
}

\forestset{
sn edges/.style={for tree={edge={-Latex}}},
  visible on/.style={
    for tree={
      /tikz/visible on={#1},
      edge+={/tikz/visible on={#1}}}}
}

\newcommand{\cmark}{\ding{51}}%

\newcommand{\xx}{\textcolor{color1}{x}}
\newcommand{\yy}{\textcolor{color3}{y}}
\newcommand{\pP}{\texttt{P}}
\newcommand{\qQ}{\texttt{Q}}
\newcommand{\dDelta}{\textcolor{color2}{\Delta}}

\newcommand{\pone}{\textcolor{color1}{\pP(1)}}
\newcommand{\ptwo}{\textcolor{color1}{\pP(2)}}
\newcommand{\qone}{\textcolor{color2}{\qQ(1)}}
\newcommand{\qtwo}{\textcolor{color2}{\qQ(2)}}
\newcommand{\pqone}{\textcolor{color3}{\pP(1)}, \textcolor{color3}{\qQ(1)}}
\newcommand{\pqtwo}{\textcolor{color3}{\pP(2)}, \textcolor{color3}{\qQ(2)}}

\newcommand{\WMC}{\textup{WMC}}
\newcommand{\WMI}{\textup{WMI}}
\newcommand{\WFOMI}{\textup{WFOMI}}
\newcommand{\AMC}{\textup{AMC}}
\newcommand{\pbp}{\textup{PBP}}
\newcommand{\SPPP}{\textup{SP}}

% 15 min (8 slides, currently have 7) introduction
% 25 min (13 slides, currently have 10) presentation
% 5 min (3 slides, currently have 1) summary
% make sure there's a clear and coherent storyline

\author{Paulius Dilkas \\ \vspace{1cm} \tiny{Joint work with Vaishak Belle (Univeristy of Edinburgh, UK)}}
\title{Synthesising Recursive Functions for\\ First-Order Model Counting}
\institute{National University of Singapore, Singapore}
\date{\today}
% TODO: update the date
% TODO: unify the use of colours
% TODO: do I want to mention forclift and/or crane?
% TODO: maybe replace red!50 with one of the main colours
% TODO: \color -> \textcolor everywhere
% TODO: how do we capitalise blocks?

\begin{document}
\addtobeamertemplate{block begin}{\setlength\abovedisplayskip{0pt}}

\begin{frame}[noframenumbering,plain]
  \tikz[remember picture,overlay]{
    \node at ([yshift=20pt,xshift=-20pt]current page.south)
    {\includegraphics[height=40pt]{nus.jpg}};
    \node at ([yshift=25pt,xshift=30pt]current page.south)
    {\includegraphics[height=40pt]{inf.png}};
    \node at ([yshift=25pt,xshift=75pt]current page.south)
    {\includegraphics[height=40pt]{ecr.jpg}};
    \node at ([yshift=20pt,xshift=140pt]current page.south)
    {\includegraphics[height=20pt]{epsrc.png}};
  }
  \titlepage
\end{frame}

\begin{frame}{What Computers Can and Cannot Do}
  \begin{tikzpicture}[remember picture,overlay]
    \node[right=0.7cm] at (current page.west) (stickman) {\includegraphics[height=0.5\textheight]{stickman.png}};
    \node<-4>[left,shift=({0, -2})] at (current page.east) (computer) {\includegraphics[height=0.5\textheight]{happy_computer}};
    \node<2>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.5)}] at (stickman.north east) {$\int_{-\infty}^{\infty} e^{-t^{2}}\,\mathrm{d}t$};
    \node<2>[ellipse callout, draw, left=-0.5cm, below=0.5cm] at (computer.north west) {$\sqrt{\pi}$};
    \node<3>[ellipse callout, draw, right=-1cm, callout relative pointer={(-2.5, -0.5)}] at (stickman.north east) {Produce a schedule for the nurses at NUH.};
    \node<3>[ellipse callout, draw, below=0.5cm, shift=({-0.5,0}), callout relative pointer={(0.5, -0.25)}] at (computer.north west) {\includegraphics[width=0.1\textwidth]{schedule}};
    \node<4>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)}] at (stickman.north east) {Describe this picture. \includegraphics[width=0.25\textwidth]{cat.jpg}};
    \node<4>[ellipse callout, draw, below=0.5cm, shift=({-0.5,0.5}), callout relative pointer={(0.5, -0.5)}] at (computer.north west) {It's a cat with its eyes closed.};
    \node<5>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)}, text width=5cm] at (stickman.north east) {If I shuffle a deck of \alert{$n$} cards, how many possible outcomes are there?};
    \node<5>[left,shift=({-1, -1})] at (current page.east) (computer2) {\includegraphics[height=0.5\textheight]{overheating_computer}\blfootnote{Terms and conditions apply.}};
    \node[shift=({1.3, 0.1})] at (current page.south west) {\tiny \color{lightgray}{Vector art by Vecteezy.com}};
  \end{tikzpicture}
\end{frame}
% TODO: replace the cat thing with a funny DALL-E picture

\begin{frame}{Who Cares About Counting?}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \pause
      \begin{block}{\alert{Probabilistic Programming}}
        \centering
        \includegraphics[width=\textwidth]{programming.png}
      \end{block}
      \pause
      \begin{block}{\alert{Neuro-symbolic AI}}
        \centering
        \includegraphics[width=\textwidth]{neurosymbolic.png}
      \end{block}
      \pause
      \begin{block}{\alert{Natural Language Processing}}
        \centering
        \includegraphics[width=\textwidth]{knowledge.png}
      \end{block}
      \pause
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{block}{\alert{Robotics}}
        \centering
        \includegraphics[width=\textwidth]{ICRA.png}
      \end{block}
      \pause
      \begin{block}{\alert{Bioinformatics}}
        \centering
        \includegraphics[width=\textwidth]{bio.png}
      \end{block}
      \pause
      \begin{block}{\alert{Combinatorics}}
        \centering
        \includegraphics[width=\textwidth]{recurrence.png}
      \end{block}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]{(Some of the) Many Ways to Count}
  \begin{center}
    \begin{tikzcd}[remember picture]
      |[alias=WMI,visible on=<7->]|\WMI{} & |[visible on=<7->]|\WFOMI{} \\
      |[alias=WMC]|\#\SAT{}/\WMC{} \arrow[u,visible on=<7->] \ar[d,visible on=<8->] \ar[r,visible on=<5->] & |[visible on=<5->]|\textup{(W)\alert<9>{FOMC}} \arrow[u,visible on=<7->] \ar[d,visible on=<8->] \\
      |[alias=PBP,visible on=<8->]|\AMC{}, \pbp{} & |[alias=SP,visible on=<8->]|\SPPP{}
    \end{tikzcd}
  \end{center}
  \begin{tikzpicture}[overlay,remember picture]
    \begin{scope}[transform canvas={xshift=-3em}]
      \draw[->,draw=gray] ($(WMC)+(0,0.1)$)--($(WMI)+(0,0.5)$) node[midway,left,visible on=<7->] {\color{gray}{$\sum \to \int$}};
      \draw[->,draw=gray] ($(WMC)-(0,0.1)$)--($(PBP)-(0,0.5)$) node[midway,left,visible on=<8->] {\color{gray}{Generalising weights}};
    \end{scope}
    \begin{scope}[transform canvas={yshift=-2em}]
      \draw[->,draw=gray] ($(PBP)-(1,0)$)--($(SP)+(1,0)$) node[midway,below,visible on=<5->] {\color{gray}{From propositional to first-order logic}};
    \end{scope}
  \end{tikzpicture}
  \vfill
    \begin{overprint}
      \onslide<1,2>
      \begin{block}{\#SAT~\color{gray}{\parencite{DBLP:journals/tcs/Valiant79}}}
        \begin{itemize}
          \item Input formula: $\xx \lor \yy$
          \item Interpretations: $\emptyset$, $\{\, \xx \,\}$, $\{\, \yy \,\}$,
                $\{\, \xx, \yy \,\}$
          \item Models: $\{\, \xx \,\}$, $\{\, \yy \,\}$, $\{\, \xx, \yy \,\}$
          \item<2-> Answer (model count): 3
        \end{itemize}
      \end{block}
      \onslide<3,4>
      \begin{block}{Weighted Model Counting~\color{gray}{\parencite{DBLP:journals/ai/ChaviraD08}}}
        \begin{itemize}
          \item Input formula: $\xx \lor \yy$
          \item Input weights: $\begin{aligned}[t]
                                  &w(\xx) = 0.3\text{, }w(\neg \xx) = 0.7,\\
                                  &w(\yy) = 0.2\text{, }w(\neg \yy) = 0.8
                                \end{aligned}$
          \item<4-> Answer (weighted model count):
                $w(\xx)w(\yy) + w(\xx)w(\neg \yy) + w(\neg \xx)w(\yy) = 0.44$
        \end{itemize}
      \end{block}
      \onslide<5,6>
      \begin{block}{(Weighted) (Symmetric) First-Order Model Counting\\
          \color{gray}{\parencite{DBLP:conf/ijcai/BroeckTMDR11}}}
        \begin{itemize}
          \item Input formula: $\forall \xx \in \Delta\text{. }\pP(\xx)$
          \item Input weights: $w^{+}(\pP) = 0.3$, $w^{-}(\pP) = 0.7$
          \item Input domain size(s): $|\dDelta| = 2$
          \item<6-> Answer: ${(w^{+}(\pP))}^{|\dDelta|} = 0.09$
        \end{itemize}
      \end{block}
      \onslide<7>
      \begin{block}{Extensions to Continuous Domains}
        \begin{itemize}
          \item Weighted model integration
          \begin{itemize}
            \item \color{gray}{\parencite{DBLP:conf/ijcai/BellePB15}}
          \end{itemize}
          \item Weighted first-order model integration
          \begin{itemize}
            \item \color{gray}{\parencite{DBLP:conf/uai/FeldsteinB21}}
          \end{itemize}
        \end{itemize}
      \end{block}
      \onslide<8>
      \begin{block}{Generalisations of the weight function}
        \begin{itemize}
          \item Algebraic model counting
          \begin{itemize}
            \item \textcolor{gray}{\parencite{DBLP:journals/japll/KimmigBR17}}
            \item From $\mathbb{R}_{\ge 0}$ to commutative semirings
          \end{itemize}
          \item Pseudo-Boolean projection~\textcolor{gray}{\parencite{DBLP:conf/sat/DilkasB21}}
          \begin{itemize}
            \item Weights not necessarily on literals
          \end{itemize}
          \item Semiring programming~\color{gray}{\parencite{DBLP:journals/ijar/BelleR20}}
        \end{itemize}
      \end{block}
    \end{overprint}
\end{frame}
% maybe mention #CSP and other non-SAT ways of counting (SMT?, counting paths in
% automata and directed graphs)

% TODO: add a slide about other approaches to WFOMC, making it crystal clear
% what my contributions are. This will finish the introduction

\begin{frame}{(Unweighted) First-Order Model Counting}
  \begin{itemize}
    \item Example formula: \\
          \alert{$\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$}
    \item Let \alert{$\Delta \coloneqq \{\, 1, 2 \,\}$}
    \item Interpretations: all subsets of\\
          \alert{$\{\, \pP(1), \qQ(1), \pP(2), \qQ(2) \,\}$}
    \item<2-> Models:
    \[
      \begin{matrix}
        \{\, \pone, \ptwo \,\}, & \{\, \pone, \qtwo \,\}, & \{\, \pone, \pqtwo \,\},\\
        \{\, \qone, \ptwo \,\}, & \{\, \qone, \qtwo \,\}, & \{\, \qone, \pqtwo \,\},\\
        \{\, \pqone, \ptwo \,\}, & \{\, \pqone, \qtwo \,\}, & \{\, \pqone, \pqtwo \,\}\\
      \end{matrix}
    \]
  \end{itemize}
  \onslide<3->{
    \begin{block}{Intuition}
      \begin{itemize}
        \item Each 1-ary predicate is like a \alert{subset}
        \item For \alert{$n>1$}, each \alert{$n$}-ary predicate is like a
              \alert{relation}
        \item FOMC counts \alert{combinations of relations}
      \end{itemize}
    \end{block}
  }
  \begin{tikzpicture}[remember picture,overlay,visible on=<2->]
    \tikzset{shift={(current page.north east)}}
    \tikzset{shift={(-4, -2.9)}}
    \node[draw,circle,minimum size=3cm,label={135:$\pP$},fill=color1] (P) at (0,0) {};
    \node[draw,circle,minimum size=3cm,label={45:$\qQ$},fill=color2] (Q) at (1.8,0) {};
    \begin{scope}
      \clip (0,0) circle(1.5cm);
      \clip (1.8,0) circle(1.5cm);
      \fill[color3](0,0) circle(1.5cm);
    \end{scope}
    \draw (0,0) circle(1.5cm);
    \draw (1.8,0) circle(1.5cm);
  \end{tikzpicture}
\end{frame}

% \begin{frame}{First-Order Model Counting (in More Detail)}
%   \begin{itemize}
%     \item \alert{First-order model counting} (FOMC) is the problem of counting
%           the models of a sentence in first-order logic.
%     \item The \alert{(symmetric) weighted} variation of the problem adds weights
%           (e.g., probabilities) to predicates.
%           \begin{itemize}
%             \item It is used for efficient \alert{probabilistic inference} in
%                   relational models such as Markov logic networks.
%           \end{itemize}
%     \item None of the (implemented) (W)FOMC algorithms are able to count, e.g.,
%           \alert{injective} and \alert{bijective} functions.
%   \end{itemize}
%   \begin{block}{Claim}
%     This shortcoming can be addressed via support for \alert{recursive functions}.
%   \end{block}
% \end{frame}

\begin{frame}{First-Order Knowledge Compilation: An Approach to FOMC}
  \begin{minipage}[t][0.81\textheight][t]{\textwidth}
    \centering
    \begin{forest}
      for tree={s sep=10mm, sn edges}
      [\alt<3->{$\bigwedge_{c \in \Delta}$}{$\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$},
      inner on=<3->,
      label={[visible on=<{13-14,19-}>]right:{\alert{\alt<15->{$3^{|\Delta|}$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [\alt<5->{$\lor$}{$\pP(c) \lor \qQ(c)$},
      name=lor,
      for tree={visible on=<3->},
      inner on=<5->,
      label={[visible on=<{12-14,18-}>]right:{\alert{\only<18->{$2 + 1 = 3$}\only<12-13>{\colorbox{white}{$\{\,\pP(c), \qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\pP(c), \qQ(c)\,\}$}}}}}
      [\alt<7->{$\land$}{$\begin{aligned}\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=and,
      for tree={visible on=<5->},
      inner on=<7->,
      label={[visible on=<{11-14,17-}>]left:{\alert{\only<17->{$2 \times 1 = 2$}\only<11-13>{\colorbox{white}{$\{\,\qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\qQ(c)\,\}$}}}}}
      [\alt<15->{$\pP(c) \lor \neg \pP(c)$}{$\qQ(c)$},
      name=smoothing,
      for tree={visible on=<7->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{2}}}]
      [\alt<15->{$\land$}{$\top$},
      for tree={visible on=<7->},
      leaf on=<6-14>,
      inner on=<15->,
      label={[label distance=0cm,visible on=<17->]87:{\alert{$1 \times 1 = 1$}}},
      label={[visible on=<10>]below:{\alert{tautology}}},
      [$\qQ(c)$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\top$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      [\alt<9->{$\land$}{$\begin{aligned}\neg\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=negformula,
      for tree={visible on=<5->},
      inner on=<9->,
      label={[visible on=<{11-14,17-}>]right:{\alert{\alt<15->{$1 \times 1 = 1$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [$\neg \qQ(c)$,
      name=negq,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\pP(c)$,
      name=pnode,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      ]
      \begin{scope}[on background layer]
        \uncover<6>{\node[fill=redish, ellipse, fit={(and)}, inner sep=0.5pt] {};}
        \uncover<8>{\node[fill=redish, ellipse, fit={(negformula)}, inner sep=0.5pt] {};}
        \uncover<15>{\node[fill=redish, ellipse, fit={(smoothing) (and)}, inner sep=0.5pt] {};}
      \end{scope}
      \node<10>[below = 3cm of lor] (n1) {\alert{unit clauses}};
      \path[-Latex,draw]<10> (n1) edge[bend right] (pnode);
      \path[-Latex,draw]<10> (n1) edge (negq);
      \path[-Latex,draw]<10> (n1) edge[bend left] (smoothing);
    \end{forest}
  \end{minipage}
  \begin{minipage}{\textwidth}
    \begin{overlayarea}{\textwidth}{4cm}
      \only<2-3>{\centerline{Independent partial grounding (introduces a constant \alert{$c \in \Delta$})}}%
      \only<4-5>{\centerline{Shannon decomposition (a.k.a. Boole's expansion theorem) on \alert{$\qQ(c)$}}}%
      \only<6-7>{\centerline{Positive unit propagation of \alert{$\qQ(c)$}}}%
      \only<8-9>{\centerline{Negative unit propagation of \alert{$\neg\qQ(c)$}}}%
      \only<10>{\centerline{Compilation is complete \textcolor{forestgreen}{\cmark}}}%
      \only<11-13>{\centerline{Smoothing: propagating atoms upwards}}%
      \only<14-15>{\centerline{Smoothing: adding new atoms}}%
      \only<16->{\centerline{Propagating the model count}}%
    \end{overlayarea}
  \end{minipage}
\end{frame}

\begin{frame}{A (Slightly) More Complicated Example}
  Suppose this room has \alert{$n$} seats, and there are \alert{$m \le n$}
  people in the audience. How many ways are there to seat everyone?
  \vfill

  \pause
  More explicitly, we assume that:
  \begin{itemize}
  \item each attendee gets exactly one seat,
  \item and a seat can accommodate at most one person.
  \end{itemize}
  \vfill

  \pause
  \alert{Answer:} $n^{\underline{m}} = n \cdot (n-1)\cdots(n-m+1)$.

  Note: this problem is equivalent to counting \alert{$[m] \to [n]$} injections.

\end{frame}
% wouldn't it be nice if we could just describe this problem and have an algorithm solve it for us?

\begin{frame}{Let's Express This Problem in Logic!}
  \begin{itemize}
    \item Let \alert{$\Gamma$} and \alert{$\Delta$} be sets (i.e., domains)
    \begin{itemize}
      \item such that \alert{$|\Gamma| = m$}, and \alert{$|\Delta| = n$}.
    \end{itemize}
    \item Let $\pP \subseteq \Gamma \times \Delta$ be a relation (i.e.,
          predicate) over \alert{$\Gamma$} and \alert{$\Delta$}.
    \item We can describe all of the constraints in first-order logic:
    \begin{itemize}
      \item \pause each attendee gets a seat (i.e., at least one seat)
      \begin{equation}\label{eq:a1}
        \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }\pP(x, y)
      \end{equation}
      \item \pause one person cannot occupy multiple seats
      \begin{equation}\label{eq:a2}
        \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\pP(x, y) \land \pP(x, z) \Rightarrow y=z
      \end{equation}
      \item \pause one seat cannot accommodate multiple attendees
      \begin{equation}\label{eq:a3}
        \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\pP(w, y) \land \pP(x, y) \Rightarrow w=x
      \end{equation}
    \end{itemize}
  \end{itemize}
  \pause \eqref{eq:a1} and \eqref{eq:a2} constrain $\pP$ to be a function, and
  \eqref{eq:a3} makes it injective.
\end{frame}
% one could also add probabilities to these sentences to say that, e.g., there is a 1% probability of somebody taking up two seats, somebody sitting on somebody else's lap, somebody sitting on the floor, etc.

{
  \usebackgroundtemplate{\includegraphics[keepaspectratio,width=\paperwidth]{matryoshka.jpg}}
  \begin{frame}{Recursion}
  \end{frame}
}

% TODO: it would be good to have a similar compilation walkthrough for an instance with Ref as well

\begin{frame}{Back to Our Example}
  The following function counts injections:
  \[
  f(m, n) =
  \begin{cases}
    1 & \text{if } m = 0 \text{ and } n = 0 \\
    0 & \text{if } m > 0 \text{ and } n = 0 \\
    f(m, n-1) + mf(m-1, n-1) & \text{otherwise.}
  \end{cases}
  \]
  \pause
  \begin{itemize}
    \item \alert{$f(m, n)$} can be computed in \alert{$\Theta(mn)$} time
    \begin{itemize}
      \item using dynamic programming
    \end{itemize}\pause
    \item \alert{Optimal} time complexity to compute \alert{$n^{\underline{m}}$}
          is \alert{$\Theta(m)$}
    \item But \alert{$\Theta(mn)$} is still much better than translating to
          propositional logic and running a WMC algorithm \pause
    \item The rest of this talk is about how to find such functions automatically
  \end{itemize}
\end{frame}

\begin{frame}{First-Order Knowledge Compilation: Before and After}
  \begin{columns}[b]
    \begin{column}{0.4\textwidth}
      \centering
      \begin{tikzpicture}
        \node (start) {\tiny $\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.7cm of start] (kc) {Compilation};
        \node[below=0.7cm of kc] (graph) {
          \scalebox{.5}{
            \begin{forest}
              for tree={sn edges,fill=red!50,ellipse,draw}
              [[
              [
              []
              []
              ]
              [
              [
              []
              []
              ]
              []
              ]
              ]]
            \end{forest}
          }
        };
        \node[draw,rounded rectangle,fill=color2!50,below=0.7cm of graph] (evaluation) {Propagation};
        \node[below=0.7cm of evaluation] (end) {\includegraphics[height=0.4cm]{81.png}};
        \node[right=0.3cm of evaluation] (sizes) {\tiny $|\Delta| = 4$};
        \draw[-Latex,color=color2] (start) -- (kc);
        \draw[-Latex,color=color2] (kc) -- (graph);
        \draw[-Latex,color=color2] (graph) -- (evaluation);
        \draw[-Latex,color=color2] (evaluation) -- (end);
        \draw[-Latex,color=color2] (sizes) -- (evaluation);
      \end{tikzpicture}
    \end{column}
    \pause
    \begin{column}{0.6\textwidth}
      \centering
      \begin{tikzpicture}
        \node (start) {\tiny $\begin{gathered}
            \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }\pP(x, y)\\
            \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\pP(x, y) \land \pP(x, z) \Rightarrow y=z\\
            \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\pP(w, y) \land \pP(x, y) \Rightarrow w=x
          \end{gathered}$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.3cm of start] (kc) {Compilation};
        \node[below=0.09cm of kc] (graph) {
          \scalebox{.5}{
            \begin{forest}
              for tree={sn edges,grow=0,reversed,fill=red!50}
              [,ellipse,draw,name=gdr
              [,ellipse,draw
              [,ellipse,draw
              [,ellipse,draw
              [,ellipse,draw]
              [,ellipse,draw,name=ref]
              ]
              ]
              ]
              ]
              \draw[-Latex,bend left=20] (ref) to (gdr);
            \end{forest}
          }
        };
        \node[draw,rounded rectangle,fill=color2!50,below=0.1cm of graph] (interpretation) {Conversion};
        \node[below=0.3cm of interpretation] (initial) {\tiny $f(m, n) = \sum_{l=0}^m \binom{m}{l} [l<2] \times f(m-l, n-1)$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.3cm of initial] (simplification) {Simplification};
        \node[below=0.3cm of simplification] (simplified) {\tiny $f(m, n) = f(m, n-1) + mf(m-1, n-1)$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.3cm of simplified] (evaluation) {Evaluation};
        \node[below=0.3cm of evaluation] (end) {\includegraphics[height=0.4cm]{42.png}};
        \node[right=0.3cm of evaluation] (sizes) {\tiny $\begin{aligned}
                                                           f(0,0) = 1&\text{{,} }f(m, 0) = 0\\
                                                           m = 2&\text{{,} }n=7
                                                         \end{aligned}$};
        \draw[-Latex,color=color2] (start) -- (kc);
        \draw[-Latex,color=color2] (kc) -- ($(graph)+(0,0.2)$);
        \draw[-Latex,color=color2] ($(graph)+(0,-0.3)$) -- (interpretation);
        \draw[-Latex,color=color2] (interpretation) -- (initial);
        \draw[-Latex,color=color2] (initial) -- (simplification);
        \draw[-Latex,color=color2] (simplification) -- (simplified);
        \draw[-Latex,color=color2] (simplified) -- (evaluation);
        \draw[-Latex,color=color2] (evaluation) -- (end);
        \draw[-Latex,color=color2] (sizes) -- (evaluation);
      \end{tikzpicture}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Circuits vs Graphs}
  \begin{block}{Circuits
      \textcolor{gray}{\parencite{DBLP:conf/ijcai/BroeckTMDR11}}\ldots}
    \begin{itemize}
      \item \ldots extend d-DNNF circuits
            \textcolor{gray}{\parencite{DBLP:journals/jancl/Darwiche01}} for
            propositional knowledge compilation with \alert{more node types}
      \item \ldots are \alert{acyclic}.
    \end{itemize}
  \end{block}
  \pause
  \begin{block}{First-Order Computational Graphs (FCGs) are\ldots}
    directed \alert{\st{acyclic}} (weakly connected) graphs with:
    \begin{itemize}
    \item a single source,
    \item labelled nodes,
    \item and ordered outgoing edges.
    \end{itemize}
  \end{block}
\end{frame}

% TODO: add arrows that explain the meaning/definition of each node type
\begin{frame}{How to Interpret an FCG}
  \begin{columns}
    \begin{column}{0.25\textwidth}
      \centering
      \begin{tikzpicture}[every node/.style={draw,ellipse},edge from parent/.style={draw,-latex},sibling distance=12mm,level distance=15mm]
        \node[fill=red!50,fill on=<2>] (dr) {\textsc{GDR}}
        child {node[fill=red!50,fill on=<3>] {$\bigvee$}
                child {node[fill=red!50, fill on=<4>] {\textsc{CR}}
                  child {node[fill=red!50, fill on=<5>] {$\land$}
                    child {node[fill=red!50, fill on=<6>] {$\bot$}}
                    child {node[fill=red!50, fill on=<7>] (ref) {\textsc{Ref}}}
        }}};
        \draw[-latex, bend right] (ref) to (dr);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.75\textwidth}
      \begin{align*}
        f(m, n) &= \onslide<3->{\sum_{l=0}^m \binom{m}{l}} \onslide<6->{[l<2]} \onslide<5->{\times} \onslide<7->{f(m-l, n-1)}\\
        \onslide<8>{&= f(m, n-1) + mf(m-1, n-1)}
      \end{align*}
      \onslide<6>{
        \[
          [\phi] =
          \begin{cases}
            1 & \text{if } \phi \\
            0 & \text{if } \neg\phi
          \end{cases}
          \]
      }
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Compilation Rules}
  \begin{definition}
    A \alert{(compilation) rule} is a function that takes a \alert{formula} and returns a set of \alert{$(G, L)$} pairs, where
    \begin{itemize}
    \item \alert{$G$} is an FCG,
    \item and \alert{$L$} is a list of formulas.
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}{Example Rule: Independence}
  Input formula:
  \begin{gather}
    ({\color{color1} \forall x, y \in \Omega\text{. }x=y}) \land \label{eq:1} \tag{{\color{color1} 1}} \\
    ({\color{color2} \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\texttt{P}(x, y) \land \texttt{P}(x, z) \Rightarrow y=z}) \land \label{eq:2} \tag{{\color{color2} 2}} \\
    ({\color{color2} \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\texttt{P}(w, y) \land \texttt{P}(x, y) \Rightarrow w=x}) \label{eq:3} \tag{{\color{color2} 3}}
  \end{gather}
  \pause
  Only one \alert{(G, L)} pair:
  \[
  G =
  \begin{tikzpicture}[edge from parent/.style={draw,-latex},baseline=(current bounding box.center)]
    \node[draw,circle] {$\land$}
    child {node {\color{color1} $\star$}}
    child {node {\color{color2} $\star$}}
    ;
  \end{tikzpicture},
  \qquad
  L = \langle {\color{color1} \eqref{eq:1}}, {\color{color2} \eqref{eq:2} \land \eqref{eq:3}} \rangle
  \]
\end{frame}
% stars mark edges to nowhere and they are associated with the formulas in L via a certain order

\begin{frame}
  \frametitle<1>{New Rule 1: Generalised Domain Recursion}
  \frametitle<2>{New Rule 2: Constraint Removal}
  \frametitle<3>{New Rule 3: Identifying Possibilities for Recursion}
  \begin{columns}
    \begin{column}{0.15\textwidth}
      \centering
      \begin{tikzpicture}[every node/.style={draw,ellipse},edge from parent/.style={draw,-latex},sibling distance=10mm,level distance=10mm]
        \node[fill=red!50, fill on=<1>] (dr) {}
        child {node {}
          child {node[fill=red!50, fill on=<2>] {}
            child {node {}
              child {node {}}
              child {node[fill=red!50, fill on=<3>] (ref) {}}
            }}};
        \draw[-latex, bend right,alt=<3>{red}{black}] (ref) to (dr);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.85\textwidth}
      \begin{overprint}
        \onslide<1>
          \begin{example}
            Input formula:
            \[
              \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
            \]
            Output formula (with a new constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(\alert{c}, y) \lor \neg \texttt{P}(\alert{c}, z) \\
              \begin{multlined}
                \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined}
            \end{gather*}
            %% Output FCG:
            %% \begin{tikzpicture}[edge from parent/.style={draw,-latex}]
            %%   \node[draw,circle] {$\textsc{DR}$}
            %%   child {node {$\star$}}
            %%   ;
            %% \end{tikzpicture}
          \end{example}
        \onslide<2>
          \begin{example}
            Input formula (with a constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \begin{multlined}
                \forall x \in \alert{\Gamma}\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined} \\
              \begin{multlined}
                \forall w,x \in \alert{\Gamma}\text{. }\forall y \in \Delta\text{. }\alert{w \ne c} \land \alert{x \ne c} \land w \ne x \Rightarrow \\
                \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
              \end{multlined}
            \end{gather*}
            Output formula (with a new domain \alert{$\Gamma' \coloneqq \Gamma \setminus \{\, c \,\}$}):
            \begin{gather*}
              \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z) \\
              \forall w,x \in \alert{\Gamma'}\text{. }\forall y \in \Delta\text{. }w \ne x \Rightarrow \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
            \end{gather*}
          \end{example}
        \onslide<3>
          \begin{block}{Goal}
            Check if the input formula is isomorphic (up to domains) to a
            previously encountered formula.
          \end{block}
          \begin{block}{Rough Outline}
            \begin{enumerate}
              \item Consider pairs of `similar' clauses.
              \item Consider bijections between their sets of variables.
              \item Extend each such bijection to a map between sets of domains.
              \item If the bijection makes the clauses equal, and the domain map
                    is compatible with previous domain maps, move on to another
                    pair of clauses.
            \end{enumerate}
          \end{block}
      \end{overprint}
    \end{column}
  \end{columns}
\end{frame}

%% \begin{frame}{Compilation as Search}
%%   \begin{definition}
%%     A \alert{(search) state} is a tuple \structure{$(G, C, L)$}, where:
%%     \begin{itemize}
%%     \item \structure{$G$} is an FCG (or \texttt{null}),
%%     \item \structure{$C$} is a compilation cache that maps integers to sets of pairs $(\phi, v)$, where $\phi$ is a formula, and $v$ is a node of $G$ (which is used to identify opportunities for recursion),
%%     \item and \structure{$L$} is a list of formulas (that are yet to be compiled). (Note that the order is crucial!)
%%     \end{itemize}
%%   \end{definition}
%%   The search algorithm combines greedy and breadth-first search:
%% \end{frame}

\begin{frame}{Resulting Improvements to Counting Functions}
  Let \alert{$\Gamma$} and \alert{$\Delta$} be two sets with
  cardinalities \alert{$|\Gamma| = m$} and \alert{$|\Delta| = n$}.

  Our new rules enable \textsc{ForcLift} to efficiently count
  \alert{$\Gamma \to \Delta$} functions such as:
  \begin{itemize}
    \item injections in \alert{$\Theta(mn)$} time
          \begin{itemize}
            \item best: \alert{$\Theta(m)$}
          \end{itemize}
    \item partial injections in \alert{$\Theta(mn)$} time
          \begin{itemize}
            \item best: \alert{$\Theta({\min\{\, m, n \,\}}^2)$}
          \end{itemize}
          \item bijections in \alert{$\Theta(m)$} time
          \begin{itemize}
            \item \alert{optimal!}
          \end{itemize}
  \end{itemize}
\end{frame}
% TODO: mention Crane, don't forget to introduce forclift
% TODO: mention that everything used to be unliftable (by any algorithm with a publicly available implementation)
% TODO: what font do I want to use for forclift/crane? It's mentioned on a slide title as well

% TODO: maybe replace the zoom picture with some galaxy picture?
{ % all template changes are local to this group.
    \setbeamertemplate{navigation symbols}{}
    \begin{frame}<article:0>[plain]
        \begin{tikzpicture}[remember picture,overlay]
            \node[at=(current page.center)] {
                \includegraphics[keepaspectratio, height=\paperheight]{zoom.jpg}
            };
        \end{tikzpicture}
     \end{frame}
}

\begin{frame}{Summary \& Future Work}
  \begin{block}{Summary}
    The circuits hitherto used for FOMC become more powerful with:
    \begin{itemize}
    \item cycles,
    \item generalised domain recursion,
    \item and some more new compilation rules that support domain recursion.
    \end{itemize}
  \end{block}
  \begin{block}{Future Work}
    \begin{itemize}
    \item Automate:
      \begin{itemize}
      \item extracting and simplifying the definitions of functions,
      \item finding all base cases.
      \end{itemize}
    \item Open questions:
      \begin{itemize}
      \item What kind of \alert{sequences} are computable in this way?
      \item Would using a \alert{different logic} extend the capabilities of FOMC further?
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}
% TODO: for future work, come back to the 'many ways to count' slide and add
% more arrows (e.g., beyond FOL)
% TODO: also incorporate the last slide of the 4th talk
\end{document}