### 2023-04-13 doc/talks/3_long/plot.tex
First draft of the entire talk

% Created by tikzDevice version 0.12.3.1 on 2023-04-13 13:23:15
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (153.94, 95.40);
\begin{scope}
\path[clip] ( 36.11, 30.69) rectangle (148.44, 89.90);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11, 41.73) --
	(148.44, 41.73);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11, 57.34) --
	(148.44, 57.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11, 72.95) --
	(148.44, 72.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 36.11, 88.56) --
	(148.44, 88.56);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 49.55, 30.69) --
	( 49.55, 89.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 70.39, 30.69) --
	( 70.39, 89.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 91.23, 30.69) --
	( 91.23, 89.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] (112.07, 30.69) --
	(112.07, 89.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] (132.91, 30.69) --
	(132.91, 89.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11, 33.93) --
	(148.44, 33.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11, 49.54) --
	(148.44, 49.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11, 65.15) --
	(148.44, 65.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 36.11, 80.76) --
	(148.44, 80.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 39.13, 30.69) --
	( 39.13, 89.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 59.97, 30.69) --
	( 59.97, 89.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 80.81, 30.69) --
	( 80.81, 89.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.65, 30.69) --
	(101.65, 89.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (122.49, 30.69) --
	(122.49, 89.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.33, 30.69) --
	(143.33, 89.90);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 41.22, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 43.30, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 45.38, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 47.47, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 49.55, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 51.64, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 53.72, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 55.80, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 57.89, 33.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 59.97, 33.94) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 62.06, 33.94) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 64.14, 33.95) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 66.22, 33.96) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 68.31, 33.97) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 70.39, 33.98) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 72.48, 34.06) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 74.56, 34.02) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 76.64, 34.05) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 78.73, 34.09) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 80.81, 34.13) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 82.90, 34.18) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 84.98, 34.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 87.06, 34.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 89.15, 34.43) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 91.23, 34.56) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 93.32, 34.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 95.40, 34.89) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 97.48, 35.08) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 99.57, 35.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (101.65, 35.74) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (103.73, 36.04) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (105.82, 36.50) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (107.90, 37.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (109.99, 37.65) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (112.07, 38.43) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (114.15, 39.34) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (116.24, 40.34) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (118.32, 41.54) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (120.41, 42.91) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (122.49, 45.10) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (124.57, 47.55) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (126.66, 48.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (128.74, 52.11) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (130.83, 54.60) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (132.91, 58.27) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (134.99, 63.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (137.08, 67.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (139.16, 73.61) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (141.25, 79.65) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (143.33, 87.21) circle (  1.96);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 1.1pt,line join=round] ( 41.22, 34.87) --
	( 42.51, 34.52) --
	( 43.80, 34.23) --
	( 45.09, 33.99) --
	( 46.39, 33.79) --
	( 47.68, 33.64) --
	( 48.97, 33.53) --
	( 50.26, 33.45) --
	( 51.56, 33.40) --
	( 52.85, 33.38) --
	( 54.14, 33.38) --
	( 55.44, 33.40) --
	( 56.73, 33.44) --
	( 58.02, 33.49) --
	( 59.31, 33.56) --
	( 60.61, 33.63) --
	( 61.90, 33.71) --
	( 63.19, 33.79) --
	( 64.48, 33.88) --
	( 65.78, 33.96) --
	( 67.07, 34.05) --
	( 68.36, 34.13) --
	( 69.65, 34.21) --
	( 70.95, 34.28) --
	( 72.24, 34.35) --
	( 73.53, 34.42) --
	( 74.82, 34.47) --
	( 76.12, 34.52) --
	( 77.41, 34.56) --
	( 78.70, 34.60) --
	( 79.99, 34.63) --
	( 81.29, 34.65) --
	( 82.58, 34.67) --
	( 83.87, 34.68) --
	( 85.16, 34.69) --
	( 86.46, 34.70) --
	( 87.75, 34.70) --
	( 89.04, 34.71) --
	( 90.33, 34.72) --
	( 91.63, 34.74) --
	( 92.92, 34.76) --
	( 94.21, 34.79) --
	( 95.50, 34.84) --
	( 96.80, 34.89) --
	( 98.09, 34.97) --
	( 99.38, 35.07) --
	(100.67, 35.19) --
	(101.97, 35.34) --
	(103.26, 35.52) --
	(104.55, 35.73) --
	(105.85, 35.99) --
	(107.14, 36.29) --
	(108.43, 36.63) --
	(109.72, 37.03) --
	(111.02, 37.49) --
	(112.31, 38.01) --
	(113.60, 38.60) --
	(114.89, 39.26) --
	(116.19, 40.00) --
	(117.48, 40.82) --
	(118.77, 41.74) --
	(120.06, 42.75) --
	(121.36, 43.87) --
	(122.65, 45.10) --
	(123.94, 46.44) --
	(125.23, 47.91) --
	(126.53, 49.51) --
	(127.82, 51.24) --
	(129.11, 53.12) --
	(130.40, 55.16) --
	(131.70, 57.35) --
	(132.99, 59.72) --
	(134.28, 62.26) --
	(135.57, 64.99) --
	(136.87, 67.91) --
	(138.16, 71.04) --
	(139.45, 74.38) --
	(140.74, 77.94) --
	(142.04, 81.73) --
	(143.33, 85.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (153.94, 95.40);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16, 30.89) {0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16, 46.51) {250};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16, 62.12) {500};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 31.16, 77.73) {750};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (153.94, 95.40);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 39.13, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 59.97, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 80.81, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (101.65, 19.68) {30};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (122.49, 19.68) {40};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (143.33, 19.68) {50};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (153.94, 95.40);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 92.27,  7.64) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (153.94, 95.40);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08, 60.29) {Runtime (s)};
\end{scope}
\end{tikzpicture}

\end{document}