### 2023-05-25 doc/talks/3_long/talk.tex
Make the talk slightly longer

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{complexity}
\usepackage[style=authoryear]{biblatex}
\usepackage{forest}
\usepackage{soul}
\usepackage{csquotes}
\usepackage{pifont}

\usetikzlibrary{shapes}
\usetikzlibrary{positioning}
\usetikzlibrary{overlay-beamer-styles}
\usetikzlibrary{shapes.callouts}
\usetikzlibrary{cd}
\usetikzlibrary{calc}
\usetikzlibrary{backgrounds}
\usetikzlibrary{fit}
\usetikzlibrary{decorations.pathreplacing}

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

\renewcommand<>{\hl}[1]{\only#2{\beameroriginal{\hl}}{#1}}
\makeatletter
\newcommand\SoulColor{%
  \let\set@color\beamerorig@set@color
  \let\reset@color\beamerorig@reset@color}
\makeatother
\SoulColor

\newcommand\Wider[2][3em]{%
\makebox[\linewidth][c]{%
  \begin{minipage}{\dimexpr\textwidth+#1\relax}
  \raggedright#2
  \end{minipage}%
  }%
}

\beamertemplatenavigationsymbolsempty
\usetheme{metropolis}
\setbeamercolor{background canvas}{bg=white}

\addbibresource{talk.bib}

\definecolor{color1}{HTML}{1B9E77}
\definecolor{color2}{HTML}{7570b3}
\definecolor{color3}{HTML}{a6761d}
\definecolor{forestgreen}{HTML}{009B55}

\DeclareMathOperator{\GDR}{\textsc{GDR}}
\DeclareMathOperator{\CR}{\textsc{CR}}
\DeclareMathOperator{\Reff}{\textsc{Ref}}

\colorlet{redish}{color2!25}
\colorlet{greenish}{color1!25}
\colorlet{blueish}{color3!25}
\sethlcolor{redish}
\setstcolor{black}

\tikzset{
  inner/.style={
    ellipse,
    draw,
  },
  inner on/.style={alt=#1{inner}{}},
  leaf/.style={
    rectangle,
    draw,
    fill=gray!25,
  },
  leaf on/.style={alt=#1{leaf}{}},
  invisible/.style={opacity=0,text opacity=0},
  visible on/.style={alt=#1{}{invisible}},
  alt/.code args={<#1>#2#3}{%
    \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}} % \pgfkeysalso doesn't change the path
  },
}

\forestset{
sn edges/.style={for tree={edge={-Latex}}},
  visible on/.style={
    for tree={
      /tikz/visible on={#1},
      edge+={/tikz/visible on={#1}}}}
}

\newcommand{\cmark}{\ding{51}}%

\newcommand{\xx}{\textcolor{color1}{x}}
\newcommand{\yy}{\textcolor{color3}{y}}
\newcommand{\pP}{\texttt{P}}
\newcommand{\qQ}{\texttt{Q}}
\newcommand{\dDelta}{\textcolor{color2}{\Delta}}

\newcommand{\pone}{\textcolor{color1}{\pP(1)}}
\newcommand{\ptwo}{\textcolor{color1}{\pP(2)}}
\newcommand{\qone}{\textcolor{color2}{\qQ(1)}}
\newcommand{\qtwo}{\textcolor{color2}{\qQ(2)}}
\newcommand{\pqone}{\textcolor{color3}{\pP(1)}, \textcolor{color3}{\qQ(1)}}
\newcommand{\pqtwo}{\textcolor{color3}{\pP(2)}, \textcolor{color3}{\qQ(2)}}

\newcommand{\WMC}{\textup{WMC}}
\newcommand{\WMI}{\textup{WMI}}
\newcommand{\WFOMI}{\textup{WFOMI}}
\newcommand{\AMC}{\textup{AMC}}
\newcommand{\pbp}{\textup{PBP}}
\newcommand{\SPPP}{\textup{SP}}

\author{Paulius Dilkas \\ \vspace{1cm} \tiny{Joint work with Vaishak Belle (Univeristy of Edinburgh, UK)}}
\title{Synthesising Recursive Functions for\\ First-Order Model Counting}
\institute{National University of Singapore, Singapore}
\date{DTAI Seminar, 26th May 2023}

\begin{document}
\addtobeamertemplate{block begin}{\setlength\abovedisplayskip{0pt}}

\begin{frame}[noframenumbering,plain]
  \tikz[remember picture,overlay]{
    \node at ([yshift=20pt,xshift=-20pt]current page.south)
    {\includegraphics[height=40pt]{nus.jpg}};
    \node at ([yshift=25pt,xshift=30pt]current page.south)
    {\includegraphics[height=40pt]{inf.png}};
    \node at ([yshift=25pt,xshift=75pt]current page.south)
    {\includegraphics[height=40pt]{ecr.jpg}};
    \node at ([yshift=20pt,xshift=140pt]current page.south)
    {\includegraphics[height=20pt]{epsrc.png}};
  }
  \titlepage
\end{frame}

\begin{frame}{What Computers Can and Cannot Do}
  \begin{tikzpicture}[remember picture,overlay]
    \node[right=0.7cm] at (current page.west) (stickman) {\includegraphics[height=0.5\textheight]{stickman.png}};
    \node<-4>[left,shift=({0, -2})] at (current page.east) (computer) {\includegraphics[height=0.5\textheight]{happy_computer}};
    \node<2>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.5)}] at (stickman.north east) {$\int_{-\infty}^{\infty} e^{-t^{2}}\,\mathrm{d}t$};
    \node<2>[ellipse callout, draw, left=-0.5cm, below=0.5cm] at (computer.north west) {$\sqrt{\pi}$};
    \node<3>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)},text width=5cm] at (stickman.north east) {Produce a \alert{schedule} for the nurses at the local hospital.};
    \node<3>[ellipse callout, draw, below=0.5cm, shift=({-0.5,0}), callout relative pointer={(0.5, -0.25)}] at (computer.north west) {\includegraphics[width=0.1\textwidth]{schedule}};
    \node<4>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)},text width=6cm] at (stickman.north east) {Paint a baroque oil painting of a \alert{raccoon queen wearing a crown}.};
    \node<4>[ellipse callout, draw, below=0cm, shift=({-1.5,1}), callout relative pointer={(0.4, -0.2)}] at (computer.north west) {\includegraphics[width=0.25\textwidth]{dalle.png}};
    \node<5>[ellipse callout, draw, right=0cm, callout relative pointer={(-1, -0.15)}, text width=5cm] at (stickman.north east) {If I shuffle a deck of \alert{$n$} cards, how many possible outcomes are there?};
    \node<5>[left,shift=({-1, -1})] at (current page.east) (computer2) {\includegraphics[height=0.5\textheight]{overheating_computer}\blfootnote{Terms and conditions apply.}};
    \node[shift=({1.3, 0.1})] at (current page.south west) {\tiny \textcolor{lightgray}{Vector art by Vecteezy.com}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Who Cares About Counting?}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \pause
      \begin{block}{\alert{Probabilistic Programming}}
        \centering
        \includegraphics[width=\textwidth]{programming.png}
      \end{block}
      \pause
      \begin{block}{\alert{Neuro-symbolic AI}}
        \centering
        \includegraphics[width=\textwidth]{neurosymbolic.png}
      \end{block}
      \pause
      \begin{block}{\alert{Natural Language Processing}}
        \centering
        \includegraphics[width=\textwidth]{knowledge.png}
      \end{block}
      \pause
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{block}{\alert{Robotics}}
        \centering
        \includegraphics[width=\textwidth]{ICRA.png}
      \end{block}
      \pause
      \begin{block}{\alert{Bioinformatics}}
        \centering
        \includegraphics[width=\textwidth]{bio.png}
      \end{block}
      \pause
      \begin{block}{\alert{Combinatorics}}
        \centering
        \includegraphics[width=\textwidth]{recurrence.png}
      \end{block}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]{(Some of the) Many Ways to Count}
  \begin{center}
    \begin{tikzcd}[remember picture]
      |[alias=WMI,visible on=<7->]|\WMI{} & |[visible on=<7->]|\WFOMI{} \\
      |[alias=WMC]|\#\SAT{}/\WMC{} \arrow[u,visible on=<7->] \ar[d,visible on=<8->] \ar[r,visible on=<5->] & |[visible on=<5->]|\textup{(W)\alert<9>{FOMC}} \arrow[u,visible on=<7->] \ar[d,visible on=<8->] \\
      |[alias=PBP,visible on=<8->]|\AMC{}, \pbp{} & |[alias=SP,visible on=<8->]|\SPPP{}
    \end{tikzcd}
  \end{center}
  \begin{tikzpicture}[overlay,remember picture]
    \begin{scope}[transform canvas={xshift=-3em}]
      \draw[->,draw=gray] ($(WMC)+(0,0.1)$)--($(WMI)+(0,0.5)$) node[midway,left,visible on=<7->] {\textcolor{gray}{$\sum \to \int$}};
      \draw[->,draw=gray] ($(WMC)-(0,0.1)$)--($(PBP)-(0,0.5)$) node[midway,left,visible on=<8->] {\textcolor{gray}{Generalising weights}};
    \end{scope}
    \begin{scope}[transform canvas={yshift=-2em}]
      \draw[->,draw=gray] ($(PBP)-(1,0)$)--($(SP)+(1,0)$) node[midway,below,visible on=<5->] {\textcolor{gray}{From propositional to first-order logic}};
    \end{scope}
  \end{tikzpicture}
  \vfill
    \begin{overprint}
      \onslide<1,2>
      \begin{block}{\#SAT~\textcolor{gray}{\parencite{DBLP:journals/tcs/Valiant79}}}
        \begin{itemize}
          \item Input formula: $\xx \lor \yy$
          \item Interpretations: $\emptyset$, $\{\, \xx \,\}$, $\{\, \yy \,\}$,
                $\{\, \xx, \yy \,\}$
          \item Models: $\{\, \xx \,\}$, $\{\, \yy \,\}$, $\{\, \xx, \yy \,\}$
          \item<2-> Answer (model count): 3
        \end{itemize}
      \end{block}
      \onslide<3,4>
      \begin{block}{Weighted Model Counting~\textcolor{gray}{\parencite{DBLP:journals/ai/ChaviraD08}}}
        \begin{itemize}
          \item Input formula: $\xx \lor \yy$
          \item Input weights: $\begin{aligned}[t]
                                  &w(\xx) = 0.3\text{, }w(\neg \xx) = 0.7,\\
                                  &w(\yy) = 0.2\text{, }w(\neg \yy) = 0.8
                                \end{aligned}$
          \item<4-> Answer (weighted model count):
                $w(\xx)w(\yy) + w(\xx)w(\neg \yy) + w(\neg \xx)w(\yy) = 0.44$
        \end{itemize}
      \end{block}
      \onslide<5,6>
      \begin{block}{(Weighted) (Symmetric) First-Order Model Counting\\
          \textcolor{gray}{\parencite{DBLP:conf/ijcai/BroeckTMDR11}}}
        \begin{itemize}
          \item Input formula: $\forall \xx \in \Delta\text{. }\pP(\xx)$
          \item Input weights: $w^{+}(\pP) = 0.3$, $w^{-}(\pP) = 0.7$
          \item Input domain size(s): $|\dDelta| = 2$
          \item<6-> Answer: ${(w^{+}(\pP))}^{|\dDelta|} = 0.09$
        \end{itemize}
      \end{block}
      \onslide<7>
      \begin{block}{Extensions to Continuous Domains}
        \begin{itemize}
          \item Weighted model integration
          \begin{itemize}
            \item \textcolor{gray}{\parencite{DBLP:conf/ijcai/BellePB15}}
          \end{itemize}
          \item Weighted first-order model integration
          \begin{itemize}
            \item \textcolor{gray}{\parencite{DBLP:conf/uai/FeldsteinB21}}
          \end{itemize}
        \end{itemize}
      \end{block}
      \onslide<8>
      \begin{block}{Generalisations of the Weight Function}
        \begin{itemize}
          \item Algebraic model counting
          \begin{itemize}
            \item \textcolor{gray}{\parencite{DBLP:journals/japll/KimmigBR17}}
            \item From $\mathbb{R}_{\ge 0}$ to commutative semirings
          \end{itemize}
          \item Pseudo-Boolean projection~\textcolor{gray}{\parencite{DBLP:conf/sat/DilkasB21}}
          \begin{itemize}
            \item Weights not necessarily on literals
          \end{itemize}
          \item Semiring programming~\textcolor{gray}{\parencite{DBLP:journals/ijar/BelleR20}}
        \end{itemize}
      \end{block}
    \end{overprint}
\end{frame}

\begin{frame}{(Unweighted) First-Order Model Counting}
  \begin{itemize}
    \item Example formula: \\
          \alert{$\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$}.
    \item Let \alert{$\Delta \coloneqq \{\, 1, 2 \,\}$}.
    \item Interpretations: all subsets of\\
          \alert{$\{\, \pP(1), \qQ(1), \pP(2), \qQ(2) \,\}$}.
    \item<2-> Models:
    \[
      \begin{matrix}
        \{\, \pone, \ptwo \,\}, & \{\, \pone, \qtwo \,\}, & \{\, \pone, \pqtwo \,\},\\
        \{\, \qone, \ptwo \,\}, & \{\, \qone, \qtwo \,\}, & \{\, \qone, \pqtwo \,\},\\
        \{\, \pqone, \ptwo \,\}, & \{\, \pqone, \qtwo \,\}, & \{\, \pqone, \pqtwo \,\}.\\
      \end{matrix}
    \]
  \end{itemize}
  \onslide<2->{
    \begin{block}{Intuition}
      \begin{itemize}
        \item Each 1-ary predicate is like a \alert{subset}.
        \item For \alert{$n>1$}, each \alert{$n$}-ary predicate is like a
              \alert{relation}.
        \item FOMC counts \alert{combinations of relations}.
      \end{itemize}
    \end{block}
  }
  \begin{tikzpicture}[remember picture,overlay,visible on=<2->]
    \tikzset{shift={(current page.north east)}}
    \tikzset{shift={(-4, -2.9)}}
    \node[draw,circle,minimum size=3cm,label={135:$\pP$},fill=color1] (P) at (0,0) {};
    \node[draw,circle,minimum size=3cm,label={45:$\qQ$},fill=color2] (Q) at (1.8,0) {};
    \begin{scope}
      \clip (0,0) circle(1.5cm);
      \clip (1.8,0) circle(1.5cm);
      \fill[color3](0,0) circle(1.5cm);
    \end{scope}
    \draw (0,0) circle(1.5cm);
    \draw (1.8,0) circle(1.5cm);
  \end{tikzpicture}
\end{frame}

\begin{frame}{More Formally: What Is the Input?}
  \[
    \forall x \in \alert<2>{\Gamma}\text{. }\forall y, z \in \alert<2>{\Delta}\text{. }\pP(x, y) \land \pP(x, z) \Rightarrow \alert<4>{y=z}
  \]
  \begin{block}{\alert<2>{Many-Sorted} \alert<3>{Function-Free} First-Order Logic with \alert<4>{Equality}}
  \begin{itemize}
    \item<5-> Any number of variables
    \item<5-> All variables are bound
    \item<6-> \alert{$\forall$} and \alert{$\exists$} quantifiers can be nested arbitrarily deep
    \item<6-> All domains are finite
    \item<6-> Predicates can have any arity
  \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Exact Algorithms for FOMC}
  \begin{itemize}
    \item \alert{ForcLift}~\textcolor{gray}{\parencite{DBLP:conf/ijcai/BroeckTMDR11}}
    \begin{itemize}
      \item knowledge compilation to \alert{FO d-DNNF}
    \end{itemize}
    \item \alert{L2C}~\textcolor{gray}{\parencite{DBLP:conf/kr/KazemiP16}}
    \begin{itemize}
      \item knowledge compilation to \alert{C++} code
    \end{itemize}
    \item \alert{Alchemy}~\textcolor{gray}{\parencite{DBLP:journals/cacm/GogateD16}}
    \begin{itemize}
      \item \alert{DPLL}-style search
    \end{itemize}
    \item \alert{FastWFOMC}~\textcolor{gray}{\parencite{DBLP:conf/uai/BremenK21}}
    \begin{itemize}
      \item knowledge compilation to \alert{sd-DNNF}
    \end{itemize}
  \end{itemize}
  \pause
  \vfill
  \begin{block}{Our Contribution}
    \centering
    \begin{tikzpicture}[align=center,node distance=0.1cm]
      \node[label=below:{ForcLift}] (forclift) {\includegraphics[width=0.2\paperwidth]{forklift.png}};
      \node[right=of forclift] (plus) {\Huge$+$};
      \node[right=of plus,label=below:{Recursion}] (recursion) {\includegraphics[width=0.2\paperwidth]{matryoshka.jpg}};
      \node[right=of recursion] (equal) {\Huge$=$};
      \node[right=of equal,label=below:{Crane}] (crane) {\includegraphics[width=0.2\paperwidth]{crane.jpg}};
    \end{tikzpicture}
  \end{block}
\end{frame}
% NOTE: explain what knowledge compilation is

\begin{frame}{ForcLift and First-Order Knowledge Compilation}
  \begin{minipage}[t][0.81\textheight][t]{\textwidth}
    \centering
    \begin{forest}
      for tree={s sep=10mm, sn edges}
      [\alt<3->{$\bigwedge_{c \in \Delta}$}{$\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$},
      inner on=<3->,
      label={[visible on=<{13-14,19-}>]right:{\alert{\alt<15->{$3^{|\Delta|}$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [\alt<5->{$\lor$}{$\pP(c) \lor \qQ(c)$},
      name=lor,
      for tree={visible on=<3->},
      inner on=<5->,
      label={[visible on=<{12-14,18-}>]right:{\alert{\only<18->{$2 + 1 = 3$}\only<12-13>{\colorbox{white}{$\{\,\pP(c), \qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\pP(c), \qQ(c)\,\}$}}}}}
      [\alt<7->{$\land$}{$\begin{aligned}\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=and,
      for tree={visible on=<5->},
      inner on=<7->,
      label={[visible on=<{11-14,17-}>]left:{\alert{\only<17->{$2 \times 1 = 2$}\only<11-13>{\colorbox{white}{$\{\,\qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\qQ(c)\,\}$}}}}}
      [\alt<15->{$\pP(c) \lor \neg \pP(c)$}{$\qQ(c)$},
      name=smoothing,
      for tree={visible on=<7->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{2}}}]
      [\alt<15->{$\land$}{$\top$},
      for tree={visible on=<7->},
      leaf on=<6-14>,
      inner on=<15->,
      label={[label distance=0cm,visible on=<17->]87:{\alert{$1 \times 1 = 1$}}},
      label={[visible on=<10>]below:{\alert{tautology}}},
      [$\qQ(c)$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\top$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      [\alt<9->{$\land$}{$\begin{aligned}\neg\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=negformula,
      for tree={visible on=<5->},
      inner on=<9->,
      label={[visible on=<{11-14,17-}>]right:{\alert{\alt<15->{$1 \times 1 = 1$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [$\neg \qQ(c)$,
      name=negq,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\pP(c)$,
      name=pnode,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      ]
      \begin{scope}[on background layer]
        \uncover<6>{\node[fill=redish, ellipse, fit={(and)}, inner sep=0.5pt] {};}
        \uncover<8>{\node[fill=redish, ellipse, fit={(negformula)}, inner sep=0.5pt] {};}
        \uncover<15>{\node[fill=redish, ellipse, fit={(smoothing) (and)}, inner sep=0.5pt] {};}
      \end{scope}
      \node<10>[below = 3cm of lor] (n1) {\alert{unit clauses}};
      \path[-Latex,draw]<10> (n1) edge[bend right] (pnode);
      \path[-Latex,draw]<10> (n1) edge (negq);
      \path[-Latex,draw]<10> (n1) edge[bend left] (smoothing);
    \end{forest}
  \end{minipage}
  \begin{minipage}{\textwidth}
    \begin{overlayarea}{\textwidth}{4cm}
      \only<2-3>{\centerline{Independent partial grounding (introduces a constant \alert{$c \in \Delta$})}}%
      \only<4-5>{\centerline{Shannon decomposition (a.k.a. Boole's expansion theorem) on \alert{$\qQ(c)$}}}%
      \only<6-7>{\centerline{Positive unit propagation of \alert{$\qQ(c)$}}}%
      \only<8-9>{\centerline{Negative unit propagation of \alert{$\neg\qQ(c)$}}}%
      \only<10>{\centerline{Compilation is complete \textcolor{forestgreen}{\cmark}}}%
      \only<11-13>{\centerline{Smoothing: propagating atoms upwards}}%
      \only<14-15>{\centerline{Smoothing: adding new atoms}}%
      \only<16->{\centerline{Propagating the model count}}%
    \end{overlayarea}
  \end{minipage}
\end{frame}

\begin{frame}{A (Slightly) More Complicated Example}
  Suppose this room has \alert{$n$} seats, and there are \alert{$m \le n$}
  people in the audience. How many ways are there to seat everyone?
  \vfill

  \pause
  More explicitly, we assume that:
  \begin{itemize}
  \item each attendee gets exactly one seat,
  \item and a seat can accommodate at most one person.
  \end{itemize}
  \vfill

  \pause
  \alert{Answer:} $n^{\underline{m}} = n \cdot (n-1)\cdots(n-m+1)$.

  Note: this problem is equivalent to counting \alert{$[m] \to [n]$} injections.

\end{frame}
% wouldn't it be nice if we could just describe this problem and have an algorithm solve it for us?

\begin{frame}{Let's Express This Problem in Logic!}
  \begin{itemize}
    \item Let \alert{$\Gamma$} and \alert{$\Delta$} be sets (i.e., domains)
    \begin{itemize}
      \item such that \alert{$|\Gamma| = m$}, and \alert{$|\Delta| = n$}.
    \end{itemize}
    \item Let \alert{$\pP \subseteq \Gamma \times \Delta$} be a relation (i.e.,
          predicate) over \alert{$\Gamma$} and \alert{$\Delta$}.
    \item We can describe all of the constraints in first-order logic:
    \begin{itemize}
      \item \pause each attendee gets a seat (i.e., at least one seat)
      \begin{equation}\label{eq:a1}
        \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }\pP(x, y)
      \end{equation}
      \item \pause one person cannot occupy multiple seats
      \begin{equation}\label{eq:a2}
        \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\pP(x, y) \land \pP(x, z) \Rightarrow y=z
      \end{equation}
      \item \pause one seat cannot accommodate multiple attendees
      \begin{equation}\label{eq:a3}
        \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\pP(w, y) \land \pP(x, y) \Rightarrow w=x
      \end{equation}
    \end{itemize}
  \end{itemize}
  \pause \alert{\eqref{eq:a1}} and \alert{\eqref{eq:a2}} constrain \alert{$\pP$}
  to be a function, and \alert{\eqref{eq:a3}} makes it injective.
\end{frame}
% one could also add probabilities to these sentences to say that, e.g., there is a 1% probability of somebody taking up two seats, somebody sitting on somebody else's lap, somebody sitting on the floor, etc.

{
  \usebackgroundtemplate{\includegraphics[keepaspectratio,width=\paperwidth]{matryoshka.jpg}}
  \begin{frame}{Recursion}
  \end{frame}
}

\begin{frame}{Back to Our Example}
  The following function counts injections:
  \[
  f(m, n) =
  \begin{cases}
    1 & \text{if } m = 0 \text{ and } n = 0 \\
    0 & \text{if } m > 0 \text{ and } n = 0 \\
    f(m, n-1) + m \times f(m-1, n-1) & \text{otherwise.}
  \end{cases}
  \]
  \pause
  \begin{itemize}
    \item \alert{$f(m, n)$} can be computed in \alert{$\Theta(mn)$} time
    \begin{itemize}
      \item using dynamic programming.
    \end{itemize}\pause
    \item \alert{Optimal} time complexity to compute \alert{$n^{\underline{m}}$}
          is \alert{$\Theta(m)$}.
    \item But \alert{$\Theta(mn)$} is still much better than translating to
          propositional logic and solving a \alert{$\#\P$-complete} problem.
          \pause
    \item The rest of this talk is about how to construct such functions
          automatically.
  \end{itemize}
\end{frame}

\begin{frame}{First-Order Knowledge Compilation: Before and After}
  \begin{columns}[b]
    \begin{column}{0.4\textwidth}
      \centering
      \begin{tikzpicture}
        \node (start) {\tiny $\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.7cm of start] (kc) {Compilation};
        \node[below=0.7cm of kc] (graph) {
          \scalebox{.5}{
            \begin{forest}
              for tree={sn edges,fill=red!50,ellipse,draw}
              [[
              [
              []
              []
              ]
              [
              [
              []
              []
              ]
              []
              ]
              ]]
            \end{forest}
          }
        };
        \node[draw,rounded rectangle,fill=color2!50,below=0.7cm of graph] (evaluation) {Propagation};
        \node[below=0.7cm of evaluation] (end) {\includegraphics[height=0.4cm]{81.png}};
        \node[right=0.3cm of evaluation] (sizes) {\tiny $|\Delta| = 4$};
        \draw[-Latex,color=color2] (start) -- (kc);
        \draw[-Latex,color=color2] (kc) -- (graph);
        \draw[-Latex,color=color2] (graph) -- (evaluation);
        \draw[-Latex,color=color2] (evaluation) -- (end);
        \draw[-Latex,color=color2] (sizes) -- (evaluation);
        \node[draw,fit={(kc) (graph) (evaluation)},inner xsep=1pt,inner ysep=2pt] {};
      \end{tikzpicture}
    \end{column}
    \pause
    \begin{column}{0.6\textwidth}
      \centering
      \begin{tikzpicture}
        \node (start) {\tiny $\begin{gathered}
            \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }\pP(x, y)\\
            \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\pP(x, y) \land \pP(x, z) \Rightarrow y=z\\
            \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\pP(w, y) \land \pP(x, y) \Rightarrow w=x
          \end{gathered}$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.3cm of start] (kc) {Compilation};
        \node[below=0.09cm of kc] (graph) {
          \scalebox{.5}{
            \begin{forest}
              for tree={sn edges,grow=0,reversed,fill=red!50}
              [,ellipse,draw,name=gdr
              [,ellipse,draw
              [,ellipse,draw
              [,ellipse,draw
              [,ellipse,draw]
              [,ellipse,draw,name=ref]
              ]
              ]
              ]
              ]
              \draw[-Latex,bend left=20] (ref) to (gdr);
            \end{forest}
          }
        };
        \node[draw,rounded rectangle,fill=color2!50,below=0.1cm of graph] (interpretation) {Conversion};
        \node[below=0.3cm of interpretation] (initial) {\tiny $f(m, n) = \sum_{l=0}^m \binom{m}{l} [l<2] \times f(m-l, n-1)$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.3cm of initial] (simplification) {Simplification};
        \node[below=0.3cm of simplification] (simplified) {\tiny $f(m, n) = f(m, n-1) + m \times f(m-1, n-1)$};
        \node[draw,rounded rectangle,fill=color2!50,below=0.3cm of simplified] (evaluation) {Evaluation};
        \node[below=0.3cm of evaluation] (end) {\includegraphics[height=0.4cm]{42.png}};
        \node[right=0.3cm of evaluation] (sizes) {\tiny $\begin{aligned}
                                                           f(0,0) = 1&\text{{,} }f(m, 0) = 0\\
                                                           m = 2&\text{{,} }n=7
                                                         \end{aligned}$};
        \draw[-Latex,color=color2] (start) -- (kc);
        \draw[-Latex,color=color2] (kc) -- ($(graph)+(0,0.2)$);
        \draw[-Latex,color=color2] ($(graph)+(0,-0.3)$) -- (interpretation);
        \draw[-Latex,color=color2] (interpretation) -- (initial);
        \draw[-Latex,color=color2] (initial) -- (simplification);
        \draw[-Latex,color=color2] (simplification) -- (simplified);
        \draw[-Latex,color=color2] (simplified) -- (evaluation);
        \draw[-Latex,color=color2] (evaluation) -- (end);
        \draw[-Latex,color=color2] (sizes) -- (evaluation);
        \node[draw,fit={(kc) (graph) (interpretation)},inner xsep=1pt,inner ysep=2pt,visible on=<3->] {};
      \end{tikzpicture}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Circuits vs Graphs}
  \begin{block}{Circuits
      \textcolor{gray}{\parencite{DBLP:conf/ijcai/BroeckTMDR11}}\ldots}
    \begin{itemize}
      \item \ldots extend d-DNNF circuits
            \textcolor{gray}{\parencite{DBLP:journals/jancl/Darwiche01}} for
            propositional knowledge compilation with \alert{more node types}
      \item \ldots are \alert{acyclic}.
    \end{itemize}
  \end{block}
  \pause
  \begin{block}{First-Order Computational Graphs (FCGs) are\ldots}
    directed \alert{\st{acyclic}} (weakly connected) graphs with:
    \begin{itemize}
    \item a single source,
    \item labelled nodes,
    \item and ordered outgoing edges.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{How to Interpret an FCG}
  \begin{columns}
    \begin{column}{0.25\textwidth}
      \centering
      \vspace{1cm}
      \begin{forest}
        for tree={sn edges}
        [$\GDR$,ellipse,draw,name=gdr,fill=red!50,fill on=<4>
        [$\bigvee$,ellipse,draw,name=bigvee,fill=red!50,fill on=<5>
        [$\CR$,ellipse,draw,name=cr,fill=red!50,fill on=<6>
        [$\land$,ellipse,draw,name=conj,fill=red!50,fill on=<7>
        [$\bot$,rectangle,draw,alt=<8>{fill=red!50}{fill=gray!25},name=contradiction]
        [$\Reff$,ellipse,draw,name=ref,fill=red!50,fill on=<9>]
        ]
        ]
        ]
        ]
        \draw[-Latex,bend right=45] (ref) to (gdr);
        \draw[visible on=<2-3>] (5, 42 |- gdr) node[text width=5cm,name=gdrtwo,text=color2] {\alert<3>{Generalised domain recursion}};
        \draw[visible on=<2-3>] (5, 42 |- bigvee) node[text width=5cm,name=bigveetwo,text=color2] {Deterministic set-disjunction};
        \draw[visible on=<2-3>] (5, 42 |- cr) node[text width=5cm,name=crtwo,text=color2] {\alert<3>{Constraint removal}};
        \draw[visible on=<2-3>] (5, 42 |- conj) node[text width=5cm,name=conjtwo,text=color2] {Decomposable conjunction};
        \draw[visible on=<2-3>] (5, 42 |- ref) node[text width=5cm,name=reftwo,text=color2] {\alert<3>{Caching}};
        \node[visible on=<2-3>,below=0.5cm of contradiction,name=contradictiontwo,text=color2] {Contradiction};
        \draw[visible on=<2-3>,-Latex,dashed,color=color2] (gdrtwo) -- (gdr);
        \draw[visible on=<2-3>,-Latex,dashed,color=color2] (bigveetwo) -- (bigvee);
        \draw[visible on=<2-3>,-Latex,dashed,color=color2] (crtwo) -- (cr);
        \draw[visible on=<2-3>,-Latex,dashed,color=color2] (conjtwo) -- (conj);
        \draw[visible on=<2-3>,-Latex,dashed,color=color2] (reftwo) -- (ref);
        \draw[visible on=<2-3>,-Latex,dashed,color=color2] (contradictiontwo) -- (contradiction);
      \end{forest}
    \end{column}
    \begin{column}{0.75\textwidth}
      \begin{align*}
        \onslide<4->{f(m, n) &=} \onslide<5->{\sum_{l=0}^m \binom{m}{l}} \onslide<8->{[l<2]} \onslide<7->{\times} \onslide<9->{f(m-l, n-1)}\\
        \onslide<10->{&= \binom{m}{0} \times f(m-0, n-1)}\\
        \onslide<10->{&+ \binom{m}{1} \times f(m-1, n-1)}\\
        \onslide<11->{&= f(m, n-1) + m \times f(m-1, n-1)}
      \end{align*}
      \onslide<8>{
        \[
          [\phi] =
          \begin{cases}
            1 & \text{if } \phi \\
            0 & \text{if } \neg\phi
          \end{cases}
          \]
      }
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Compilation: How FCGs Are Built}
  \begin{definition}
    A \alert{(compilation) rule} is a function that takes a \alert{formula} and returns a set of \alert{$(G, L)$} pairs, where
    \begin{itemize}
    \item \alert{$G$} is a (possibly incomplete) FCG,
    \item and \alert{$L$} is a list of formulas.
    \end{itemize}
  \end{definition}
  \vfill The formulas in \alert{$L$} are then \alert{compiled}, and the
  resulting FCGs are \alert{inserted} into \alert{$G$} according to a \alert{set
    order}.
\end{frame}

\begin{frame}{Example Compilation Rule: Independence}
  Input formula:
  \begin{gather}
    ({\color{color1} \forall x, y \in \Omega\text{. }x=y}) \land \label{eq:1} \tag{{\color{color1} 1}} \\
    ({\color{color2} \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\texttt{P}(x, y) \land \texttt{P}(x, z) \Rightarrow y=z}) \land \label{eq:2} \tag{{\color{color2} 2}} \\
    ({\color{color2} \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\texttt{P}(w, y) \land \texttt{P}(x, y) \Rightarrow w=x}) \label{eq:3} \tag{{\color{color2} 3}}
  \end{gather}
  \pause
  The independence compilation rule returns one \alert{(G, L)} pair:
  \[
  G =
  \begin{tikzpicture}[edge from parent/.style={draw,-Latex},baseline=(current bounding box.center)]
    \node[draw,circle] {$\land$}
    child {node {\color{color1} $\star$}}
    child {node {\color{color2} $\star$}}
    ;
  \end{tikzpicture},
  \qquad
  L = \langle {\color{color1} \eqref{eq:1}}, {\color{color2} \eqref{eq:2} \land \eqref{eq:3}} \rangle
  \]
\end{frame}
% stars mark edges to nowhere and they are associated with the formulas in L via a certain order

\begin{frame}
  \frametitle<1>{New Rule 1/3: Generalised Domain Recursion}
  \frametitle<2>{New Rule 2/3: Constraint Removal}
  \frametitle<3>{New Rule 3/3: Identifying Possibilities for Recursion}
  \begin{columns}
    \begin{column}{0.15\textwidth}
      \centering
      \begin{tikzpicture}[every node/.style={draw,ellipse},edge from parent/.style={draw,-Latex},sibling distance=10mm,level distance=10mm]
        \node[fill=red!50, fill on=<1>] (dr) {}
        child {node {}
          child {node[fill=red!50, fill on=<2>] {}
            child {node {}
              child {node {}}
              child {node[fill=red!50, fill on=<3>] (ref) {}}
            }}};
        \draw[-Latex, bend right,alt=<3>{red}{black}] (ref) to (dr);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.85\textwidth}
      \begin{overprint}
        \onslide<1>
          \begin{example}
            Input formula:
            \[
              \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
            \]
            Output formula (with a new constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(\alert{c}, y) \lor \neg \texttt{P}(\alert{c}, z) \\
              \begin{multlined}
                \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined}
            \end{gather*}
            %% Output FCG:
            %% \begin{tikzpicture}[edge from parent/.style={draw,-latex}]
            %%   \node[draw,circle] {$\textsc{DR}$}
            %%   child {node {$\star$}}
            %%   ;
            %% \end{tikzpicture}
          \end{example}
        \onslide<2>
          \begin{example}
            Input formula (with a constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \begin{multlined}
                \forall x \in \alert{\Gamma}\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined} \\
              \begin{multlined}
                \forall w,x \in \alert{\Gamma}\text{. }\forall y \in \Delta\text{. }\alert{w \ne c} \land \alert{x \ne c} \land w \ne x \Rightarrow \\
                \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
              \end{multlined}
            \end{gather*}
            Output formula (with a new domain \alert{$\Gamma' \coloneqq \Gamma \setminus \{\, c \,\}$}):
            \begin{gather*}
              \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z) \\
              \forall w,x \in \alert{\Gamma'}\text{. }\forall y \in \Delta\text{. }w \ne x \Rightarrow \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
            \end{gather*}
          \end{example}
        \onslide<3>
          \begin{block}{Goal}
            Check if the input formula is equivalent (up to domains) to a
            previously encountered formula.
          \end{block}
          \begin{block}{Outline}
            \begin{enumerate}
              \item Consider pairs of `similar' clauses.
              \item Consider bijections between their sets of variables.
              \item Extend each such bijection to a map between sets of domains.
              \item If the bijection makes the clauses equivalent, and the
                    domain map is compatible with previous domain maps, move on
                    to another pair of clauses.
            \end{enumerate}
          \end{block}
      \end{overprint}
    \end{column}
  \end{columns}
\end{frame}

%% \begin{frame}{Compilation as Search}
%%   \begin{definition}
%%     A \alert{(search) state} is a tuple \structure{$(G, C, L)$}, where:
%%     \begin{itemize}
%%     \item \structure{$G$} is an FCG (or \texttt{null}),
%%     \item \structure{$C$} is a compilation cache that maps integers to sets of pairs $(\phi, v)$, where $\phi$ is a formula, and $v$ is a node of $G$ (which is used to identify opportunities for recursion),
%%     \item and \structure{$L$} is a list of formulas (that are yet to be compiled). (Note that the order is crucial!)
%%     \end{itemize}
%%   \end{definition}
%%   The search algorithm combines greedy and breadth-first search:
%% \end{frame}

\begin{frame}{How These Rules Fit Together (1/5)}
  \begin{tikzpicture}[overlay,remember picture]
    \node[draw,rounded rectangle,fill=color2!50,anchor=center,above=0.5cm] at (current page.center) (rules) {Generalised domain recursion};
    \node[above=of rules] (before) {$\begin{gathered}
        \forall x \in \alert{\Gamma}\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z)\\
        \forall w,x \in \alert{\Gamma}\text{. }\forall y \in \Delta\text{. }w \ne x \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \node[below=of rules] (after) {$\begin{gathered}
        \forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg\pP(\alert{c}, y) \lor \neg\pP(\alert{c}, z)\\
        \forall x \in \Gamma\text{. }\forall y,z \in \Delta\text{. }y \ne z \land \alert{x \ne c} \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z)\\
        \forall x \in \Gamma\text{. }\forall y \in \Delta\text{. }\alert{x \ne c} \Rightarrow \neg\pP(\alert{c}, y) \lor \neg\pP(x, y)\\
        \forall w \in \Gamma\text{. }\forall y \in \Delta\text{. }\alert{w \ne c} \Rightarrow \neg\pP(w, y) \lor \neg\pP(\alert{c}, y)\\
        \forall w,x \in \Gamma\text{. }\forall y \in \Delta\text{. }w \ne x \land \alert{w \ne c} \land \alert{x \ne c} \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \begin{scope}[on background layer]
      \node[anchor=north,fill=greenish,minimum width=11.1cm,minimum height=0.7cm] at (before.north) {};
      \node[anchor=south,fill=blueish,minimum width=11.1cm,minimum height=0.7cm] at (before.south) {};
      \node[anchor=north,fill=greenish,minimum width=11.1cm,minimum height=1.3cm] at (after.north) {};
      \node[anchor=south,fill=blueish,minimum width=11.1cm,minimum height=2cm] at (after.south) {};
    \end{scope}
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (before) -- (rules);
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (rules) -- (after);
  \end{tikzpicture}
\end{frame}

\begin{frame}{How These Rules Fit Together (2/5)}
  \begin{tikzpicture}[overlay,remember picture]
    \node[draw,rounded rectangle,fill=color2!50,anchor=center,below=1cm] at (current page.center) (rules) {Atom counting and unit propagation};
    \node[above=of rules] (before) {$\begin{gathered}
        \forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg\pP(\alert{c}, y) \lor \neg\pP(\alert{c}, z)\\
        \forall x \in \Gamma\text{. }\forall y,z \in \Delta\text{. }y \ne z \land \alert{x \ne c} \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z)\\
        \forall x \in \Gamma\text{. }\forall y \in \Delta\text{. }\alert{x \ne c} \Rightarrow \neg\pP(\alert{c}, y) \lor \neg\pP(x, y)\\
        \forall w \in \Gamma\text{. }\forall y \in \Delta\text{. }\alert{w \ne c} \Rightarrow \neg\pP(w, y) \lor \neg\pP(\alert{c}, y)\\
        \forall w,x \in \Gamma\text{. }\forall y \in \Delta\text{. }w \ne x \land \alert{w \ne c} \land \alert{x \ne c} \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \node[below=of rules] (after) {$\begin{gathered}
        \forall y, z \in \alert{\Delta^{\top}}\text{. }y \ne z \Rightarrow \bot\\
        \forall x \in \Gamma\text{. }\forall y, z \in \alert{\Delta^{\bot}}\text{. }y \ne z \land x \ne c \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z) \\
        \forall w, x \in \Gamma\text{. }\forall y \in \alert{\Delta^{\bot}}\text{. }w \ne x \land w \ne c \land x \ne c \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (before) -- (rules);
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (rules) -- (after);
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (rules) -- (after);
  \end{tikzpicture}
\end{frame}
% NOTE: need to explain how atom counting splits the domain

\begin{frame}{How These Rules Fit Together (3/5)}
  \begin{tikzpicture}[overlay,remember picture]
    \node[draw,rounded rectangle,fill=color2!50,anchor=center] at (current page.center) (rules) {Constraint removal};
    \node[above=of rules] (before) {$\begin{gathered}
        \forall y, z \in \Delta^{\top}\text{. }y \ne z \Rightarrow \bot\\
        \forall x \in \alert{\Gamma}\text{. }\forall y, z \in \Delta^{\bot}\text{. }y \ne z \land \alert{x \ne c} \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z) \\
        \forall w, x \in \alert{\Gamma}\text{. }\forall y \in \Delta^{\bot}\text{. }w \ne x \land \alert{w \ne c} \land \alert{x \ne c} \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \node[below=of rules] (after) {$\begin{gathered}
        \forall y, z \in \Delta^{\top}\text{. }y \ne z \Rightarrow \bot\\
        \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \Delta^{\bot}\text{. }y \ne z \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z) \\
        \forall w, x \in \alert{\Gamma'}\text{. }\forall y \in \Delta^{\bot}\text{. }w \ne x \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (before) -- (rules);
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (rules) -- (after);
  \end{tikzpicture}
\end{frame}

\begin{frame}{How These Rules Fit Together (4/5)}
  \begin{tikzpicture}[overlay,remember picture]
    \node[draw,rounded rectangle,fill=color2!50,anchor=center] at (current page.center) (rules) {Independence};
    \node[above=of rules] (before) {$\begin{gathered}
        \forall y, z \in \alert{\Delta^{\top}}\text{. }y \ne z \Rightarrow \bot\\
        \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \alert{\Delta^{\bot}}\text{. }y \ne z \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z) \\
        \forall w, x \in \alert{\Gamma'}\text{. }\forall y \in \alert{\Delta^{\bot}}\text{. }w \ne x \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \node[below left=of rules,fill=greenish] (after1) {$\forall y, z \in \alert{\Delta^{\top}}\text{. }y \ne z \Rightarrow \bot$};
    \node[below right=2cm and -4cm of rules,fill=blueish] (after2) {$\begin{gathered}
        \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \alert{\Delta^{\bot}}\text{. }y \ne z \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z) \\
        \forall w, x \in \alert{\Gamma'}\text{. }\forall y \in \alert{\Delta^{\bot}}\text{. }w \ne x \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \begin{scope}[on background layer]
      \node[anchor=north,fill=greenish,minimum width=9cm,minimum height=0.7cm] at (before.north) {};
      \node[anchor=south,fill=blueish,minimum width=9cm,minimum height=1.3cm] at (before.south) {};
    \end{scope}
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (before) -- (rules);
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (rules) -- (after1);
    \draw[-Latex,color=color2,ultra thick,shorten <=0.1cm,shorten >=0.1cm] (rules) -- (after2);
  \end{tikzpicture}
\end{frame}

\begin{frame}{How These Rules Fit Together (5/5): Recursion}
  \centering
  \begin{tikzpicture}[overlay,remember picture]
    \node[anchor=center,above=0.5cm] at (current page.center) (equal) {\rotatebox{90}{\textcolor{color2}{\Huge$=$}}};
    \node[below=0.5cm of equal] (after1) {$\begin{gathered}
        \forall x \in \alert{\Gamma}\text{. }\forall y, z \in \alert{\Delta}\text{. }y \ne z \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z)\\
        \forall w,x \in \alert{\Gamma}\text{. }\forall y \in \alert{\Delta}\text{. }w \ne x \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \node[above=0.5cm of equal] (foo) {$\begin{gathered}
        \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \alert{\Delta^{\bot}}\text{. }y \ne z \Rightarrow \neg\pP(x, y) \lor \neg\pP(x, z) \\
        \forall w, x \in \alert{\Gamma'}\text{. }\forall y \in \alert{\Delta^{\bot}}\text{. }w \ne x \Rightarrow \neg\pP(w, y) \lor \neg\pP(x, y)
      \end{gathered}$};
    \node[below=0.5cm of after1] (plus) {\textcolor{color2}{\Huge$+$}};
    \node[below=0.5cm of plus] {$\{\, \alert{\Gamma} \mapsto \alert{\Gamma'}, \alert{\Delta} \mapsto \alert{\Delta^{\bot}} \,\}$};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Resulting Improvements to Counting Functions}
  \Wider[3em]{ Let \alert{$\Gamma$} and \alert{$\Delta$} be two sets with
    cardinalities \alert{$|\Gamma| = m$} and \alert{$|\Delta| = n$}.

    Our new rules enable Crane to efficiently count \alert{$\Gamma \to \Delta$}
    functions such as: \vspace{1em}
    \begin{itemize}
            \setlength\itemsep{1em}
      \item injections in \alert{$\Theta(mn)$} time
            \begin{itemize}
              \item by hand: \alert{$\Theta(m)$}
            \end{itemize}
      \item partial injections in \alert{$\Theta(mn)$} time
            \begin{itemize}
              \item by hand: \alert{$\Theta({\min\{\, m, n \,\}}^2)$}
            \end{itemize}
      \item bijections in \alert{$\Theta(m)$} time
            \begin{itemize}
              \item \alert{optimal!}\pause
              \item In comparison, FastWFOMC scales as \alert{$\Omega(m^{4})$}.
            \end{itemize}
    \end{itemize}
  }
  \begin{tikzpicture}[remember picture,overlay]
    \node[draw,rectangle,color=gray] at ([xshift=-3.3cm,yshift=-0.2cm]current page.east) {\includegraphics{plot}};
  \end{tikzpicture}
\end{frame}

{ % all template changes are local to this group.
    \setbeamertemplate{navigation symbols}{}
    \begin{frame}<article:0>[plain]
        \begin{tikzpicture}[remember picture,overlay]
            \node[at=(current page.center)] {
                \includegraphics[keepaspectratio, width=\paperwidth]{hubble.jpg}
            };
            \node[shift=({-0.7, 0.1})] at (current page.south east) {\tiny \textcolor{lightgray}{Source: NASA}};
        \end{tikzpicture}
     \end{frame}
}

\begin{frame}{What Have We Learned?}
  \begin{itemize}
    \item Knowledge compilation can build \alert{graphs with cycles}.
    \item Graphs (as well as circuits) define \alert{functions}.
    \item Cycles can represent \alert{recursive calls}, including:
    \begin{itemize}
      \item mutual recursion
      \item and function calls as complex as \alert{$f(n-k-2)$}.
    \end{itemize}
    \item Recursion helps us solve counting problems that were \alert{previously
          beyond the reach of FOMC}\@.
    \item In some cases, even if a polynomial-time solution is already known,
          Crane is able to find \alert{more efficient solutions}, with a lower
          degree polynomial.
  \end{itemize}
\end{frame}

{ % all template changes are local to this group.
    \setbeamertemplate{navigation symbols}{}
    \begin{frame}<article:0>[plain]
        \begin{tikzpicture}[remember picture,overlay]
            \node[at=(current page.center)] {
                \includegraphics[keepaspectratio, height=\paperheight]{future.jpg}
            };
        \end{tikzpicture}
     \end{frame}
}

\begin{frame}[fragile]{Beyond First-Order Logic}
  \vspace{0.5cm}
  \hspace*{0.25\linewidth}
  \begin{tikzcd}[remember picture]
    |[alias=WMI]|\WMI{} & \WFOMI{} & \\
    |[alias=WMC]|\#\SAT{}/\WMC{} \arrow[u] \ar[d] \ar[r] & \textup{(W)FOMC} \arrow[u] \ar[d] \ar[r,dashed] & \raisebox{-0.4\height}{\includegraphics[width=0.1\linewidth]{question-mark.jpg}} \\
    |[alias=PBP]|\AMC{}, \pbp{} & |[alias=SP]|\SPPP{} &
  \end{tikzcd}
  \begin{tikzpicture}[overlay,remember picture]
    \begin{scope}[transform canvas={xshift=-3em}]
      \draw[->,draw=gray] ($(WMC)+(0,0.1)$)--($(WMI)+(0,0.5)$) node[midway,left] {\textcolor{gray}{$\sum \to \int$}};
      \draw[->,draw=gray] ($(WMC)-(0,0.1)$)--($(PBP)-(0,0.5)$) node[midway,left] {\textcolor{gray}{Generalising weights}};
    \end{scope}
    \begin{scope}[transform canvas={yshift=-2em}]
      \draw[->,draw=gray] ($(PBP)-(1,0)$)--($(SP)+(1,0)$) node[midway,below] {\textcolor{gray}{From propositional to first-order logic}};
    \end{scope}
  \end{tikzpicture}
  \vfill
  \vspace{1.5cm}
  What kind of logic is needed to succinctly describe, e.g.,
  \begin{itemize}
    \item $f(n) = f(f(n-1))$
    \item or the Fibonacci sequence?
  \end{itemize}
\end{frame}
% impossibility results as well (finite model theory)

\begin{frame}{Algebraic Solutions to Parameterised Problems (1/2)}
  \begin{itemize}
    \item Suppose we have a Markov logic network that models the probability
          \alert{$P$} that some system will fail.
    \item Here:
    \begin{itemize}
      \item domain sizes describe the numbers of various components,
      \item and weights express probabilities that:
      \begin{itemize}
        \item some component fails,
        \item or some combination of failures leads to another failure.
      \end{itemize}
    \end{itemize}
    \item Crane can express \alert{$P$} as a function of the domain sizes and
          weights.
  \end{itemize}
\end{frame}

\begin{frame}{Algebraic Solutions to Parameterised Problems (2/2)}
  With the help of a computer algebra system, we can then:
  \begin{itemize}
    \item determine how \alert{$P$} scales with the number of users,
    \item find combinations of domain sizes that keep \alert{$P$} below some
          threshold,
    \item find ranges of weights that keep \alert{$P$} sufficiently small across
          a range of domain size values.
  \end{itemize}
  \pause
  \vfill
  \centering
  \begin{tikzpicture}
    \node[draw] (a) {Reasoning with \alert{functions}};
    \node[right=0.5cm of a] (b) {\Huge\textcolor{color2}{$>$}};
    \node[draw,right=0.5cm of b] (c) {Reasoning with \alert{numbers}};
  \end{tikzpicture}
\end{frame}

\end{document}